---

title : (Big) Data Analytics for Business
subtitle : Session 2-3, Dimensionality Reduction and Derived Attributes
author : T. Evgeniou and J. Niessing
job : INSEAD
widgets : []
mode : standalone 

---

## What is Dimensionality Deduction and Factor Analysis?

Derive new data variables which are linear combinations of the original ones and capture most of the information in the data. 

Is typically used as a first step in Data Analytics

Can also be used to solve multicollinearity issues in regression.

---

## Factor Analysis: Key idea


1. Transform the original variables (ONLY the x's) into a smaller set of factors

2. Understand the underlying structure of the data and the new factors

3. Use the factors for subsequent analysis


```{r Fig1, fig.width=12, fig.height=6, warning=FALSE, echo=FALSE, fig.align='center', fig=TRUE}
img <- readPNG("All3.png")
grid.raster(img, vjust=0)
```

---

## Key Questions

1. Can we really simplify the data by grouping the columns?

2. How many factors should we use?

3. How good are the factors we found?

4. How interpretable are the factors we found? 

---

## Dimensionality Reduction and Factor Analysis: 8 (Easy) Steps

1. Confirm the data are metric (interval scale) 

2. Decide whether to scale standardize the data

3. Check the correlation matrix to see if Factor Analysis makes sense

4. Develop a scree plot and decide on the number of factors to be derived

5. Interpret the factors and Consider rotation of factors (technical but useful)

6. Save factor scores for subsequent analyses

---

## Applying Factor Analysis: Evaluating MBA Applications

Variables available:
* GPA
* GMAT score
* Scholarships, fellowships won
* Evidence of Communications skills 
* Prior Job Experience
* Organizational Experience
* Other extra curricular achievements

Which variables do you believe correlate with  each other?
What do these variables capture?

---

## Step 1: Confirm the data are metric (interval scale) 

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
print(xtable(summary(ProjectData) ,caption=paste("Summary Statistics:",data_name,sep=" "), digits=3), type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = FALSE)
```

---

## Step 2: Decide whether to scale standardize the data

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
ProjectData_scaled=apply(ProjectData,2, function(r) {if (sd(r)!=0) res=(r-mean(r))/sd(r) else res=0*r; res})
print(xtable(summary(ProjectData_scaled) ,caption=paste("Summary Statistics of Normalized Data:",data_name,sep=" "), digits=3), type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = FALSE)
```

---

## Step 3:  Check correlation matrix to see if Factor Analysis makes sense

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
correl<-corstars(ProjectData)
## printing the result in a clean-slate table
print(xtable(correl ,caption=paste("Correlation Matrix:",data_name,sep=" "), digits=3), type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)
```
** = correlation is significant at 1% level; * = correlation is significant at 5% level

#### Even if the data is not as neatly correlated as here, Factor analysis can be helpful

---

## Step 4. Develop a scree plot and decide on the number of factors to be derived

* Factors
* If there are n variables we will have n factors in total
* First factor will explain most variance, second next and so on.

* Variance Explained by Factors
* with standardized variables each variable has a variance of 1, so the total variance in n variables is n
* each factor will have an associated eigen value which is the amount of variance explained by that factor

--- 

## Example Factors

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
Unrotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate="none")
Unrotated_Factors<-Unrotated_Results$loadings
Unrotated_Factors<-as.data.frame(unclass(Unrotated_Factors))
colnames(Unrotated_Factors)<-paste("Component",1:ncol(Unrotated_Factors),sep=" ")
Unrotated_Factors[abs(Unrotated_Factors) < MIN_VALUE]<-NA
print(xtable(Unrotated_Factors, caption=paste(paste("All PCA factors available (values less than",MIN_VALUE),"are not shown)"), digits=3), type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE,include.colnames = TRUE)
```

--- 

## How Many Factors? Eigenvalues and Variance Explained

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
## print result in a clean-slate table
#Variance_Explained_Table[abs(Variance_Explained_Table)<MIN_VALUE]<-NA
print(xtable(Variance_Explained_Table, caption="Eigenvalues and Variance Explained", digits=3), type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)
```

--- 

## How Many Factors? Scree Plot

```{r Fig2, echo=FALSE, comment=NA, results='asis', message=FALSE, echo=FALSE, fig.align='center', fig=TRUE}
eigenvalues<-Unrotated_Results$values
## plot eigenvalues
plot(eigenvalues, type="o", main="Scree Plot",xlab="Component Number", ylab="Eigenvalue", col=colors()[26], lwd=2)
## add minor ticks
minor.tick(nx=5, ny=2)
## add abline at y = 1
abline(h=1,lty=2,col=colors()[34]) 
```

--- 

## How many factors?

* The maximal number of factors is the number of original variables Eigenvalue>1
* "Elbow" in the Scree plot
* Cumulative variance explained

---

## Step 5. Interpret the factors and Consider rotation of factors (technical but useful)

Rotated Factors

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
Rotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate=rotation_used,score=TRUE)
Rotated_Factors<-Rotated_Results$loadings
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Component",1:ncol(Rotated_Factors),sep=" ")

Rotated_Factors[abs(Rotated_Factors) < MIN_VALUE]<-NA

print(xtable(Rotated_Factors, caption=paste(paste(paste("All PCA factors available with Rotation",rotation_used),"(values less than",MIN_VALUE),"are not shown)"), digits=3), type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE,include.colnames = TRUE)
```

---

## What Factor Loads "Look Good"? Three Quality Criteria

1. For each factor (column) only a few loadings are large (in absolute value)

2. For each initial variable (row) only a few loadings are large (in absolute value)

3. Any pair of factors (columns) should have different "patterns" of loading

---

## Step 6. Save factor scores for subsequent analyses


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
NEW_ProjectData <- Rotated_Results$scores[,1:sum(eigenvalues>=1), drop=F]
colnames(NEW_ProjectData)<-paste("Derived Variable (Factor)",1:ncol(NEW_ProjectData),sep=" ")
print(xtable(head(NEW_ProjectData,5) ,paste("Sample Data described using the selected Factors:",data_name,sep=" "), digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)
```

---

## Using the Factor Scores: Pereptual Maps

```{r Fig3, echo=FALSE, comment=NA, results='asis', message=FALSE, echo=FALSE, fig.align='center', fig=TRUE}
if (ncol(NEW_ProjectData)>=2){
  plot(NEW_ProjectData[,1],NEW_ProjectData[,2], main="Data Visualization Using the top 2 Derived Attributes (Factors)",xlab="Derived Variable (Factor) 1", ylab="Derived Variable (Factor) 2", col=colors()[26], lwd=2) 
  } else {
    plot(NEW_ProjectData[,1],ProjectData[,2], 
         main="Only 1 Derived Variable: Using Initial Variable",
         xlab="Derived Variable (Factor) 1", 
         ylab="Initial Variable (Factor) 2")    
    }
```


## Factor Analysis: Key (Technical) Notions 

1. Correlation
2. Variance explained (eigenvalues)
3. Scree plot
4. Varimax rotation
5. Factor Loadings ("components")
6. Factor scores

---

## Next Class: Cluster Analysis for Segmentation

```{r include=FALSE}
## using dummy data to recreate cluster segmentation graphics
library(vegan)
require(vegan)
data(dune)
# kmeans
kclus <- kmeans(dune,centers= 4, iter.max=1000)
# distance matrix
dune_dist <- dist(dune,method="euclidian")
# Multidimensional scaling
cmd <- cmdscale(dune_dist)
```

```{r fig.width=12, fig.height=6, message=FALSE, echo=FALSE, fig.align='center', warning=FALSE, fig=TRUE}
# plot MDS, with colors by groups from kmeans
groups <- levels(factor(kclus$cluster))
ordiplot(cmd, type = "n")
cols <- c("steelblue", "darkred", "darkgreen", "pink")
for(i in seq_along(groups)){
  points(cmd[factor(kclus$cluster) == groups[i], ], col = cols[i], pch = 16)
  }

# add spider and hull
ordispider(cmd, factor(kclus$cluster), label = TRUE)
ordihull(cmd, factor(kclus$cluster), lty = "dotted")
```
