---
title: "DSB_Group1_FinalProjectProposal"
author: "Marcelo De Rada Ocampo, Ollie Phillpot, Miguel Lucas, Prathamesh Dole, Harshul Lilani"
output:
  html_document:
    css: ../AnalyticsStyles/default.css
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    includes:
      in_header: ../AnalyticsStyles/default.sty
always_allow_html: yes
---

```{r echo=FALSE, message=FALSE}
make_pdf_file = 0 # SET THIS TO 1 IF WE COMPILE PDF FILE, 0 OTHERWISE (FOR HTML)

source("../AnalyticsLibraries/library.R")
source("../AnalyticsLibraries/heatmapOutput.R")

# Package options
ggthemr('fresh')  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.2)
options(knitr.kable.NA = '')
```

``` {r echo=FALSE, message=FALSE}
# HEADS UP: This entire section has been copy pasted from the file ClassificationProcessCreditCardDefault from class
# First we will enter the filename
datafile_name = "../DATA/HousingData_Train.csv"
ProjectData <- read.csv(datafile_name)
# We turn the data into data.matrix class so that we can easier manipulate it
ProjectData <- data.matrix(ProjectData)

# Please ENTER the dependent variable (class).
# Please use numbers, not column names. E.g., 82 uses the 82nd column as the dependent variable.
# You need to make sure that dependent variable takes only two values: 0 and 1.
#dependent_variable = 25

# Please ENTER the attributes to use as independent variables. 
# Please use numbers, not column names. E.g., c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8.
#independent_variables = c(1:24) # use all the available attributes

#dependent_variable = unique(sapply(dependent_variable,function(i) min(ncol(ProjectData), max(i,1))))
#independent_variables = unique(sapply(independent_variables,function(i) min(ncol(ProjectData), max(i,1))))

#if (length(unique(ProjectData[,dependent_variable])) !=2){
 # cat("\n*****\n BE CAREFUL, THE DEPENDENT VARIABLE TAKES MORE THAN 2 VALUES")
 # cat("\nSplitting it around its median...\n*****\n ")
 # new_dependent = ProjectData[,dependent_variable] >= median(ProjectData[,dependent_variable])
 # ProjectData[,dependent_variable] <- 1*new_dependent
#}

# Please ENTER the probability threshold above which an observation is predicted as class 1:
#Probability_Threshold = 0.5 # between 0 and 1

# Please ENTER the percentage of data used for estimation
#estimation_data_percent = 80
#validation_data_percent = 10
#test_data_percent = 100-estimation_data_percent-validation_data_percent

# Please ENTER 1 if you want to randomly split the data in estimation and validation/test
#random_sampling = 0

# Tree parameter
# Please ENTER the tree (CART) complexity control cp (e.g. 0.0001 to 0.02, depending on the data)
#CART_cp = 0.0025
#CART_control = rpart.control(cp = CART_cp)

# Please ENTER the words for the business interpretation of class 1 and class 0:
#class_1_interpretation = "default"
#class_0_interpretation = "no default"

# Please ENTER the profit/cost values for correctly classified and misclassified data:
#actual_1_predict_1 = 0
#actual_1_predict_0 = -100000
#actual_0_predict_1 = 0
#actual_0_predict_0 = 20000

#Profit_Matrix = matrix(c(actual_1_predict_1, actual_0_predict_1, actual_1_predict_0, actual_0_predict_0), ncol=2)
#colnames(Profit_Matrix) <- c(paste("Predict 1 (", class_1_interpretation, ")", sep = ""), paste("Predict 0 (", #class_0_interpretation, ")", sep = ""))
#rownames(Profit_Matrix) <- c(paste("Actual 1 (", class_1_interpretation, ")", sep = ""), paste("Actual 0 (", #class_0_interpretation, ")", sep = ""))

# Please ENTER the maximum number of observations to show in the report and slides 
# (DEFAULT is 50. If the number is large the report and slides may not be generated - very slow or will crash!!)
#max_data_report = 10 
```

# The Business Context
Generally the largest investment of a person's life is buying a house. It is an emotional affair and people often overpay. There are also many small businesses that build and sell residential housing, but the construction industry is exceedingly slow to adopt new technological practices. When buying or selling property it is considered advantageous to know the area to develop a "feel" for sale prices. This is unscientific and we believe there is much room for optimisation.

We will participate in this Kaggle competion for a personal reason and a business reason:
1. Miguel is currently looking to buy a house and would like to know potential hidden factors he should look for in order to find a for a good price.
2. Ollie is a shareholder in his father's housing development company, which has recently completed a project and is looking for its next investment opportunity. This project will be used to identify possible features that could add significant value to the next project and improve ROI.

The Kaggle competition can be found [here](https://www.kaggle.com/c/house-prices-advanced-regression-techniques).

<hr>\clearpage

# The Data
(Data source: https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data?select=train.csv. We acknowledge the following:
DeCock, Dean. (2011). UCI Machine Learning Repository [http://jse.amstat.org/v19n3/decock.pdf]. Ames, Iowa: Alternative to the Boston Housing Data as an End of Semester Regression Project at Truman State University.)

The data set has been generated as an alternative for data scientists looking for a modernized and expanded version of the often cited Boston Housing dataset. The following is an example of the data/ data library:

With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa, this competition challenges participants to predict the final price of each home.

Name                       | Description
:--------------------------|:--------------------------------------------------------------------
MSSubClass | Identifies the type of dwelling involved in the sale.	
       
        20	1-STORY 1946 & NEWER ALL STYLES
        30	1-STORY 1945 & OLDER
        40	1-STORY W/FINISHED ATTIC ALL AGES
        45	1-1/2 STORY - UNFINISHED ALL AGES
        50	1-1/2 STORY FINISHED ALL AGES
        60	2-STORY 1946 & NEWER
        70	2-STORY 1945 & OLDER
        75	2-1/2 STORY ALL AGES
        80	SPLIT OR MULTI-LEVEL
        85	SPLIT FOYER
        90	DUPLEX - ALL STYLES AND AGES
       120	1-STORY PUD (Planned Unit Development) - 1946 & NEWER
       150	1-1/2 STORY PUD - ALL AGES
       160	2-STORY PUD - 1946 & NEWER
       180	PUD - MULTILEVEL - INCL SPLIT LEV/FOYER
       190	2 FAMILY CONVERSION - ALL STYLES AND AGES

MSZoning | Identifies the general zoning classification of the sale.
       
       A	Agriculture
       C	Commercial
       FV	Floating Village Residential
       I	Industrial
       RH	Residential High Density
       RL	Residential Low Density
       RP	Residential Low Density Park 
       RM	Residential Medium Density
	
LotFrontage | Linear feet of street connected to property

LotArea | Lot size in square feet

Street |  Type of road access to property
      
       Grvl	Gravel	
       Pave	Paved
       	
Alley | Type of alley access to property
      
       Grvl	Gravel
       Pave	Paved
       NA 	No alley access

LotShape | General shape of property
       
       Reg	Regular	
       IR1	Slightly irregular
       IR2	Moderately Irregular
       IR3	Irregular
       
LandContour | Flatness of the property
       
       Lvl	Near Flat/Level	
       Bnk	Banked - Quick and significant rise from street grade to building
       HLS	Hillside - Significant slope from side to side
       Low	Depression
		
Utilities | Type of utilities available
       
       AllPub	All public Utilities (E,G,W,& S)	
       NoSewr	Electricity, Gas, and Water (Septic Tank)
       NoSeWa	Electricity and Gas Only
       ELO	Electricity only	
	
LotConfig | Lot configuration
       
       Inside	Inside lot
       Corner	Corner lot
       CulDSac	Cul-de-sac
       FR2	Frontage on 2 sides of property
       FR3	Frontage on 3 sides of property

LandSlope | Slope of property
      
       Gtl	Gentle slope
       Mod	Moderate Slope	
       Sev	Severe Slope
	
Neighborhood | Physical locations within Ames city limits
      
       Blmngtn	Bloomington Heights
       Blueste	Bluestem
       BrDale	Briardale
       BrkSide	Brookside
       ClearCr	Clear Creek
       CollgCr	College Creek
       Crawfor	Crawford
       Edwards	Edwards
       Gilbert	Gilbert
       IDOTRR	Iowa DOT and Rail Road
       MeadowV	Meadow Village
       Mitchel	Mitchell
       Names	North Ames
       NoRidge	Northridge
       NPkVill	Northpark Villa
       NridgHt	Northridge Heights
       NWAmes	Northwest Ames
       OldTown	Old Town
       SWISU	South & West of Iowa State University
       Sawyer	Sawyer
       SawyerW	Sawyer West
       Somerst	Somerset
       StoneBr	Stone Brook
       Timber	Timberland
       Veenker	Veenker
			
Condition1 | Proximity to various conditions
       
       Artery	Adjacent to arterial street
       Feedr	Adjacent to feeder street	
       Norm	Normal	
       RRNn	Within 200' of North-South Railroad
       RRAn	Adjacent to North-South Railroad
       PosN	Near positive off-site feature--park, greenbelt, etc.
       PosA	Adjacent to postive off-site feature
       RRNe	Within 200' of East-West Railroad
       RRAe	Adjacent to East-West Railroad
	
Condition2 | Proximity to various conditions (if more than one is present)
       
       Artery	Adjacent to arterial street
       Feedr	Adjacent to feeder street	
       Norm	Normal	
       RRNn	Within 200' of North-South Railroad
       RRAn	Adjacent to North-South Railroad
       PosN	Near positive off-site feature--park, greenbelt, etc.
       PosA	Adjacent to postive off-site feature
       RRNe	Within 200' of East-West Railroad
       RRAe	Adjacent to East-West Railroad
	
BldgType | Type of dwelling
      
       1Fam	Single-family Detached	
       2FmCon	Two-family Conversion; originally built as one-family dwelling
       Duplx	Duplex
       TwnhsE	Townhouse End Unit
       TwnhsI	Townhouse Inside Unit
	
HouseStyle | Style of dwelling
      
       1Story	One story
       1.5Fin	One and one-half story: 2nd level finished
       1.5Unf	One and one-half story: 2nd level unfinished
       2Story	Two story
       2.5Fin	Two and one-half story: 2nd level finished
       2.5Unf	Two and one-half story: 2nd level unfinished
       SFoyer	Split Foyer
       SLvl	Split Level
	
OverallQual | Rates the overall material and finish of the house
     
       10	Very Excellent
       9	Excellent
       8	Very Good
       7	Good
       6	Above Average
       5	Average
       4	Below Average
       3	Fair
       2	Poor
       1	Very Poor
	
OverallCond | Rates the overall condition of the house
       
       10	Very Excellent
       9	Excellent
       8	Very Good
       7	Good
       6	Above Average	
       5	Average
       4	Below Average	
       3	Fair
       2	Poor
       1	Very Poor
		
YearBuilt | Original construction date

YearRemodAdd | Remodel date (same as construction date if no remodeling or additions)

RoofStyle | Type of roof
      
       Flat	Flat
       Gable	Gable
       Gambrel	Gabrel (Barn)
       Hip	Hip
       Mansard	Mansard
       Shed	Shed
		
RoofMatl | Roof material
     
       ClyTile	Clay or Tile
       CompShg	Standard (Composite) Shingle
       Membran	Membrane
       Metal	Metal
       Roll	Roll
       Tar&Grv	Gravel & Tar
       WdShake	Wood Shakes
       WdShngl	Wood Shingles
		
Exterior1st | Exterior covering on house
     
       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast	
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles
	
Exterior2nd | Exterior covering on house (if more than one material)
     
       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles
	
MasVnrType | Masonry veneer type
    
       BrkCmn	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       None	None
       Stone	Stone
	
MasVnrArea | Masonry veneer area in square feet

ExterQual | Evaluates the quality of the material on the exterior 
     
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       Po	Poor
		
ExterCond | Evaluates the present condition of the material on the exterior
    
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       Po	Poor
		
Foundation | Type of foundation
    
       BrkTil	Brick & Tile
       CBlock	Cinder Block
       PConc	Poured Contrete	
       Slab	Slab
       Stone	Stone
       Wood	Wood
		
BsmtQual | Evaluates the height of the basement
     
       Ex	Excellent (100+ inches)	
       Gd	Good (90-99 inches)
       TA	Typical (80-89 inches)
       Fa	Fair (70-79 inches)
       Po	Poor (<70 inches
       NA	No Basement
		
BsmtCond | Evaluates the general condition of the basement
     
       Ex	Excellent
       Gd	Good
       TA	Typical - slight dampness allowed
       Fa	Fair - dampness or some cracking or settling
       Po	Poor - Severe cracking, settling, or wetness
       NA	No Basement
	
BsmtExposure | Refers to walkout or garden level walls
     
       Gd	Good Exposure
       Av	Average Exposure (split levels or foyers typically score average or above)	
       Mn	Mimimum Exposure
       No	No Exposure
       NA	No Basement
	
BsmtFinType1 | Rating of basement finished area
    
       GLQ	Good Living Quarters
       ALQ	Average Living Quarters
       BLQ	Below Average Living Quarters	
       Rec	Average Rec Room
       LwQ	Low Quality
       Unf	Unfinshed
       NA	No Basement
		
BsmtFinSF1 | Type 1 finished square feet

BsmtFinType2 | Rating of basement finished area (if multiple types)
    
       GLQ	Good Living Quarters
       ALQ	Average Living Quarters
       BLQ	Below Average Living Quarters	
       Rec	Average Rec Room
       LwQ	Low Quality
       Unf	Unfinshed
       NA	No Basement

BsmtFinSF2 | Type 2 finished square feet

BsmtUnfSF | Unfinished square feet of basement area

TotalBsmtSF | Total square feet of basement area

Heating | Type of heating
    
       Floor	Floor Furnace
       GasA	Gas forced warm air furnace
       GasW	Gas hot water or steam heat
       Grav	Gravity furnace	
       OthW	Hot water or steam heat other than gas
       Wall	Wall furnace
		
HeatingQC | Heating quality and condition
   
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       Po	Poor
		
CentralAir | Central air conditioning
    
       N	No
       Y	Yes
		
Electrical | Electrical system
    
       SBrkr	Standard Circuit Breakers & Romex
       FuseA	Fuse Box over 60 AMP and all Romex wiring (Average)	
       FuseF	60 AMP Fuse Box and mostly Romex wiring (Fair)
       FuseP	60 AMP Fuse Box and mostly knob & tube wiring (poor)
       Mix	Mixed
		
1stFlrSF | First Floor square feet
 
2ndFlrSF | Second floor square feet

LowQualFinSF | Low quality finished square feet (all floors)

GrLivArea | Above grade (ground) living area square feet

BsmtFullBath | Basement full bathrooms

BsmtHalfBath | Basement half bathrooms

FullBath | Full bathrooms above grade

HalfBath | Half baths above grade

Bedroom | Bedrooms above grade (does NOT include basement bedrooms)

Kitchen | Kitchens above grade

KitchenQual | Kitchen quality
       
       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       	
TotRmsAbvGrd | Total rooms above grade (does not include bathrooms)

Functional | Home functionality (Assume typical unless deductions are warranted)
      
       Typ	Typical Functionality
       Min1	Minor Deductions 1
       Min2	Minor Deductions 2
       Mod	Moderate Deductions
       Maj1	Major Deductions 1
       Maj2	Major Deductions 2
       Sev	Severely Damaged
       Sal	Salvage only
		
Fireplaces | Number of fireplaces

FireplaceQu | Fireplace quality
      
       Ex	Excellent - Exceptional Masonry Fireplace
       Gd	Good - Masonry Fireplace in main level
       TA	Average - Prefabricated Fireplace in main living area or Masonry Fireplace in basement
       Fa	Fair - Prefabricated Fireplace in basement
       Po	Poor - Ben Franklin Stove
       NA	No Fireplace
		
GarageType | Garage location
      
       2Types	More than one type of garage
       Attchd	Attached to home
       Basment	Basement Garage
       BuiltIn	Built-In (Garage part of house - typically has room above garage)
       CarPort	Car Port
       Detchd	Detached from home
       NA	No Garage
		
GarageYrBlt | Year garage was built
		
GarageFinish | Interior finish of the garage
     
       Fin	Finished
       RFn	Rough Finished	
       Unf	Unfinished
       NA	No Garage
		
GarageCars | Size of garage in car capacity

GarageArea | Size of garage in square feet

GarageQual | Garage quality
    
       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       NA	No Garage
		
GarageCond | Garage condition
    
       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       NA	No Garage
		
PavedDrive | Paved driveway
     
       Y	Paved 
       P	Partial Pavement
       N	Dirt/Gravel
		
WoodDeckSF | Wood deck area in square feet

OpenPorchSF | Open porch area in square feet

EnclosedPorch | Enclosed porch area in square feet

3SsnPorch | Three season porch area in square feet

ScreenPorch | Screen porch area in square feet

PoolArea | Pool area in square feet

PoolQC | Pool quality
     
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       NA	No Pool
		
Fence | Fence quality
     
       GdPrv	Good Privacy
       MnPrv	Minimum Privacy
       GdWo	Good Wood
       MnWw	Minimum Wood/Wire
       NA	No Fence
	
MiscFeature | Miscellaneous feature not covered in other categories
    
       Elev	Elevator
       Gar2	2nd Garage (if not described in garage section)
       Othr	Other
       Shed	Shed (over 100 SF)
       TenC	Tennis Court
       NA	None
		
MiscVal | $Value of miscellaneous feature

MoSold | Month Sold (MM)

YrSold | Year Sold (YYYY)

SaleType | Type of sale
   
       WD 	Warranty Deed - Conventional
       CWD	Warranty Deed - Cash
       VWD	Warranty Deed - VA Loan
       New	Home just constructed and sold
       COD	Court Officer Deed/Estate
       Con	Contract 15% Down payment regular terms
       ConLw	Contract Low Down payment and low interest
       ConLI	Contract Low Interest
       ConLD	Contract Low Down
       Oth	Other
		
SaleCondition | Condition of sale
   
       Normal	Normal Sale
       Abnorml	Abnormal Sale -  trade, foreclosure, short sale
       AdjLand	Adjoining Land Purchase
       Alloca	Allocation - two linked properties with separate deeds, typically condo with a garage unit	
       Family	Sale between family members
       Partial	Home was not completed when last assessed (associated with New Homes)


Let's look into the data for a few customers. This is how the first r min(max_data_report, nrow(ProjectData)) out of the total of r nrow(ProjectData) rows look like (transposed, for convenience):


```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
#knitr::kable({
#  df <- t(head(round(ProjectData[,independent_variables],2), max_data_report))
#  colnames(df) <- sprintf("%02d", 1:ncol(df))
#  df
# })
```

# A Process for Classification

> It is important to remember that data analytics projects require a delicate balance between experimentation, intuition, and following a process. The value of following a process is so as to avoid getting fooled by randomness in data and finding "results and patterns" that are mainly driven by our own biases and not by the facts/data themselves.

*There is no single best process* for classification. However, we have to start somewhere, so we will use the following process:

1. Create an estimation sample and two validation samples by splitting the data into three groups. Steps 2-5 below will then be performed only on the estimation and the first validation data. You should only do step 6 once on the second validation data, also called **test data**, and only report/use the performance on that (second validation) data to make final business decisions. 
2. Set up the dependent variable (as a categorical 0-1 variable; multi-class classification is also feasible, and similar, but we do not explore it in this note). 
3. Make a preliminary assessment of the relative importance of the explanatory variables using visualization tools and simple descriptive statistics. 
4. Estimate the classification model using the estimation data, and interpret the results.
5. Assess the accuracy of classification in the first validation sample, possibly repeating steps 2-5 a few times changing the classifier in different ways to increase performance.
6. Finally, assess the accuracy of classification in the second validation sample. You should eventually use and report all relevant performance measures and plots on this second validation sample only.

Let's follow these steps.


## Step 1: Split the data 
It is very important that you (or the data scientists working on the project) finally measure and report the performance of the models on **data that have not been used at all during the analysis, called "out-of-sample" or test data** (steps 2-5 above). The idea is that in practice we want our models to be used for predicting the class of observations/data we have not seen yet (i.e., "the future data"): although the performance of a classification method may be high in the data used to estimate the model parameters, it may be significantly poorer on data not used for parameter estimation, such as the **out-of-sample** (future) data. 

This is why we split the data into an estimation sample and two validation samples  - using some kind of randomized splitting technique. The second validation data mimic out-of-sample data, and the performance on this validation set is a better approximation of the performance one should expect in practice from the selected classification method. The estimation data and the first validation data are used during steps 2-5 (with a few iterations of these steps), while the second validation data is only used once at the very end before making final business decisions based on the analysis. The split can be, for example, 80% estimation, 10% validation, and 10% test data, depending on the number of observations - for example, when there is a lot of data, you may only keep a few hundreds of them for the validation and test sets, and use the rest for estimation. 

While setting up the estimation and validation samples, you should also check that the same proportion of data from each class (i.e., customers who default versus not) are maintained in each sample. That is, you should maintain the same balance of the dependent variable categories as in the overall dataset. 

For simplicity, in this note we will not iterate steps 2-5. In practice, however, we should usually iterate steps 2-5 a number of times using the first validation sample each time, and at the end make our final assessment of the classification model using the test sample only once.

## Step 2: Set up the dependent variable
First, make sure the dependent variable is set up as a categorical 0-1 variable. In our illustrative example, we use the payment default (or no default) as the dependent variable. 

## Step 3: Simple Analysis
Good data analytics start with good contextual knowledge as well as a simple statistical and visual exploration of the data. In the case of classification, one can explore "simple classifications" by assessing how the classes differ along any of the independent variables. For example, these are the statistics of our independent variables across the two classes in the estimation data, class 1 ("default"):

## Step 4: Classification and Interpretation
Once we decide which  dependent and independent variables to use (which can be revisited in later iterations), one can use a number of classification methods to develop a model that discriminates the different classes.

> Some of the widely used classification methods are:  classification and regression trees (CART), boosted trees, support vector machines, neural networks, nearest neighbors, logistic regression, lasso, random forests, deep learning methods, etc.

In this note we will consider for simplicity only two classification methods: **logistic regression** and **classification and regression trees (CART)**. However, replacing them with other methods is relatively simple (although some knowledge of how these methods work is often necessary - see the R help command for the methods if needed). Understanding how these methods work is beyond the scope of this note - there are many references available online for all these classification methods. 

**Logistic Regression**: Logistic Regression is a method similar to linear regression except that the dependent variable is discrete (e.g., 0 or 1). **Linear** logistic regression estimates the coefficients of a linear model using the selected independent variables while optimizing a classification criterion. For example, this is the logistic regression parameters for our data:

Given a set of independent variables, the output of the estimated logistic regression (the sum of the products of the independent variables with the corresponding regression coefficients) can be used to assess the probability an observation belongs to one of the classes. Specifically, the regression output can be transformed into a probability of belonging to, say, class 1 for each observation. The estimated probability that a validation observation belongs to class 1 (e.g., the estimated probability that the customer defaults) for the first few validation observations, using the logistic regression above, is:

## Step 5: Validation accuracy
Using the predicted class probabilities of the validation data, as outlined above, we can  generate some measures of classification performance. Before discussing them, note that given the probability an observation belongs to a class, **a reasonable class prediction choice is to predict the class that has the highest probability**. However, this does not need to be the only choice in practice.

> Selecting the probability threshold based on which we predict the class of an observation is a decision the user needs to make. While in some cases a reasonable probability threshold is 50%, in other cases it may be 99.9% or 0.1%.

### 1.  Hit ratio
This is the percentage of the observations that have been correctly classified (i.e., the predicted class and the actual class are the same). We can just count the number of the validation data correctly classified and divide this number with the total number of the validation data, using the two CART and the logistic regression above. These are as follows for probability threshold :

### 2. Confusion matrix
The confusion matrix shows for each class the number (or percentage) of the  data that are correctly classified for that class. For example, for the method above with the highest hit rate in the validation data (among logistic regression and the 2 CART models), and for probability threshold , the confusion matrix for the validation data is:

### 3. ROC curve
Remember that each observation is classified by our model according to the probabilities Pr(0) and Pr(1) and a chosen probability threshold. Typically we set the probability threshold to 0.5 - so that observations for which Pr(1) > 0.5 are classified as 1's. However, we can vary this threshold, for example if we are interested in correctly predicting all 1's but do not mind missing some 0's (and vice-versa).

When we change the probability threshold we get different values of hit rate, false positive and false negative rates, or any other performance metric. We can plot for example how the false positive versus true positive rates change as we alter the probability threshold, and generate the so called ROC curve. 

The ROC curves for the validation data for the logistic regression as well as both the CARTs above are as follows:

### 4. Gains chart
The gains chart is a popular technique in certain applications, such as direct marketing or credit risk. 

For a concrete example, consider the case of a direct marketing mailing campaign. Say we have a classifier that attempts to identify the likely responders by assigning each case a probability of response. We may want to select as few cases as possible and still capture the maximum number of responders possible. 

We can measure the percentage of all responses the classifier captures if we only select, say, x% of cases: the top x% in terms of the probability of response assigned by our classifier. For each percentage of cases we select (x), we can plot the following point: the x-coordinate will be the percentage of all cases that were selected, while the y-coordinate will be the percentage of all class 1 cases that were captured within the selected cases (i.e., the ratio true positives/positives of the classifier, assuming the classifier predicts class 1 for all the selected cases, and predicts class 0 for all the remaining cases). If we plot these points while we change the percentage of cases we select (x) (i.e., while we change the probability threshold of the classifier), we get a chart that is called the **gains chart**. 

In the credit card default case we are studying, the gains charts for the validation data for our three classifiers are the following:

## Step 6. Test Accuracy
Having iterated steps 2-5 until we are satisfied with the performance of our selected model on the validation data, in this step the performance analysis outlined in step 5 needs to be done with the test sample. This is the performance that best mimics what one should expect in practice upon deployment of the classification solution, **assuming (as always) that the data used for this performance analysis are representative of the situation in which the solution will be deployed.** 

Let's see in our case how the **hit ratio, confusion matrix, ROC curve, gains chart, and profit curve** look like for our test data. For the hit ratio and the confusion matrix we use the price of the house as the probability threshold for classification.

