---
title: "Derived Attributes and Dimensionality Reduction: Interactive Tool"
author: "T. Evgeniou"
runtime: shiny
output: 
  html_document:
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes

---

```{r echo=FALSE, message=FALSE}
source("../../AnalyticsLibraries/library.R")

opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center")
options(knitr.kable.NA = '')

tags$style(type="text/css", "p { text-align:justify; }")
tags$style(type="text/css", "label { display: none; }")
tags$style(type="text/css", ".c3 svg { font-size:13px; font-family:\"Roboto\", \"Helvetica Neue\", Helvetica, Arial, sans-serif; }")
tags$style(type="text/css", ".formattable_widget { overflow:auto; max-height:425px; margin-bottom:23px; }")
tags$style(type="text/css", ".formattable_widget table { margin-bottom:0; }")
tags$style(type="text/css", ".formattable_widget td, .formattable_widget th { white-space: nowrap; }")

MAX_VARIABLES <- 30     # limits initial dataset filters
SAMPLE_ROWS <- 10       # limits sample dataset presentations
SAMPLE_XPOINTS <- 100   # limits sample dataset plots

normalize.abs <- function(x, min=0, max=1, na.rm=FALSE) normalize(abs(x), min, max, na.rm)
dformat <- function(df) {
  if (class(df) != "data.frame")
    df <- as.data.frame(df)
  x <- lapply(colnames(df), function(col) {
    if (is.numeric(df[, col]))
      color_bar(rgb(238, 238, 238, max=255), normalize.abs, min=0.1, na.rm=TRUE)
    else
      formatter("span")
  })
  names(x) <- colnames(df)
  formattable(df, x)
}
```


## Factor Analysis in 6 steps

This tool follows the 6 steps for factor analysis outlined in the [Derived Attributes and Dimensionality Reduction](http://inseaddataanalytics.github.io/INSEADAnalytics/Report_s23.html) reading of the course. 

First we load the data (`data/MBAadmin.csv` by default):

```{r}
# Please ENTER the name of the file with the data used. The file should contain a matrix with one row per observation (e.g. person) and one column per attribute.

fileInput("dataFile", "",
          accept = c(
            "text/csv",
            "text/comma-separated-values,text/plain",
            ".csv"),
          width = "100%"
          )
ProjectDataX <- eventReactive(input$dataFile, ignoreNULL=FALSE, {
  datapath <- input$dataFile$datapath
  if (is.null(datapath))
    datapath <- "data/MBAadmin.csv"
  df <- try(read.csv(datapath, stringsAsFactors=FALSE), silent=TRUE)
  df.valid <- class(df) == "data.frame" && any(sapply(df, is.numeric))
  if (!df.valid) {
    showModal(modalDialog(title="Error", HTML("The data file contains no valid numeric variables.<br><br>Please ensure it can be processed with `read.csv` in the command line. Common problems include a column separator other than the comma (`,`) and a decimal point other than the period (`.`).")))
    shiny::validate(need(df.valid, message=FALSE))
  }
  if (is.null(rownames(df)) || identical(rownames(df), as.character(1:nrow(df))))
    rownames(df) <- paste0("observation ", sprintf("%02i", 1:nrow(df)))
  df
})
```

We also need to select which variables to consider for factor analysis (`r MAX_VARIABLES` at most): 

```{r}
# Please ENTER then original raw attributes to use. 

selectizeInput("factorAttributes", "",
               choices = NULL,
               multiple = TRUE,
               options = list(
                 plugins = c("remove_button","drag_drop"),
                 maxItems = MAX_VARIABLES,
                 placeholder = "None"
               ),
               width="100%"
               )
observeEvent(ProjectDataX(), {
  attributes <- colnames(ProjectDataX())
  attributes <- attributes[sapply(ProjectDataX(), is.numeric)]
  updateSelectizeInput(session, "factorAttributes", choices=attributes, selected=head(attributes, MAX_VARIABLES))
})

ProjectDataFactorX <- eventReactive(input$factorAttributes, {
  ProjectDataX()[, input$factorAttributes, drop=F]
})
```

Here is how the first `r SAMPLE_ROWS` rows look like:

```{r}
renderFormattable(dformat(head(ProjectDataFactorX(), SAMPLE_ROWS)))
```


### Step 1: Confirm data is metric

The data we use here have the following descriptive statistics: 

```{r}
renderFormattable(dformat(round(my_summary(ProjectDataFactorX()), 2)))
```


### Step 2: Scale the  data

Select variables to standardize:

```{r}
# Please enter the variables you would like to standardize: 

selectizeInput("standardizedVariables", "",
               choices = NULL,
               multiple = TRUE,
               options = list(
                 plugins = c("remove_button","drag_drop"),
                 placeholder = "None"
               ),
               width="100%"
               )
observeEvent(ProjectDataFactorX(), {
  attributes <- colnames(ProjectDataFactorX())
  updateSelectizeInput(session, "standardizedVariables", choices=attributes, selected=attributes)
})

ProjectDataFactorStandardized <- eventReactive(input$standardizedVariables, ignoreNULL=FALSE, {
  df <- ProjectDataFactorX()
  columns <- input$standardizedVariables
  if (length(columns) == 0)
    return(df)
  df[, columns] <- apply(df[, columns, drop=F], 2, function(r) {
    if (sd(r) != 0) 
      (r - mean(r)) / sd(r)
    else
      0*r
  })
  df
})
```

These are the summary statistics of the scaled dataset:

```{r}
renderFormattable(dformat(round(my_summary(ProjectDataFactorStandardized()), 2)))
```


### Step 3:  Check correlations 

This is the correlation matrix of the original variables we use for factor analysis:

```{r}
renderFormattable(dformat(round(cor(ProjectDataFactorStandardized()), 2)))
```


### Step 4: Choose number of factors 

This is the Variance Explained table, using all data selected for factor analysis:

```{r}
Variance_Explained_TableX <- eventReactive(ProjectDataFactorStandardized(), {
  Variance_Explained_Table_results<-PCA(ProjectDataFactorStandardized(), graph=FALSE)
  Variance_Explained_Table<-Variance_Explained_Table_results$eig
  
  rownames(Variance_Explained_Table) <- paste("Component", 1:nrow(Variance_Explained_Table))
  colnames(Variance_Explained_Table) <- c("Eigenvalue", "Pct of explained variance", "Cumulative pct of explained variance")
  Variance_Explained_Table
})
renderFormattable(dformat(round(Variance_Explained_TableX(), 2)))
```

Here is the **scree plot**:

```{r}
tags$div(style='height:480px', renderC3({
  eigenvalues  <- Variance_Explained_TableX()[, "Eigenvalue"]
  df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
  colnames(df) <- c("eigenvalues", "components", "abline")

  c3(melt(df, id="components"), x="components", y="value", group="variable") %>% c3_line('spline')
}))
```

We now select the criterion to use for deciding how many factors to use:

```{r}
# Please ENTER the selection criterions for the factors to use. 
# Choices: "eigenvalue", "variance", "manual"
selectizeInput("factor_selectionciterion", "",
               choices = c("eigenvalue", "variance", "manual"))

# Please ENTER the desired minumum variance explained 
# (ONLY USED in case "variance" is the factor selection criterion used). 
conditionalPanel("input.factor_selectionciterion == 'variance'",
  tags$p("Since you selected the \"variance\" criterion, you need to specify the percentage of the variance you would like to explain with the final factors:"),
  sliderInput("minimum_variance_explained", "", min=1, max=100, value=65)
)

# Please ENTER the number of factors to use 
# (ONLY USED in case "manual" is the factor selection criterion used).
conditionalPanel("input.factor_selectionciterion == 'manual'",
  tags$p("Since you selected the number of factors manually, you need to specify how many you would like to use:"),
  sliderInput("manual_numb_factors_used", "", min=1, max=10, value=2)
)
observeEvent(Variance_Explained_TableX(), {
  updateSliderInput(session, "manual_numb_factors_used", max=nrow(Variance_Explained_TableX()))
})

factors_selectedX <- eventReactive({ 
  input$factor_selectionciterion
  input$minimum_variance_explained
  input$manual_numb_factors_used
  Variance_Explained_TableX()
}, {
  switch(input$factor_selectionciterion,
    eigenvalue = {
      sum(Variance_Explained_TableX()[, "Eigenvalue"] >= 1)
    },
    variance = {
      1:head(which(Variance_Explained_TableX()[, "Cumulative pct of explained variance"] >= input$minimum_variance_explained), 1)
    },
    manual = {
      input$manual_numb_factors_used
    }
  )
})
```


### Step 5: Interpret the factors

We can now use a rotation to get easier to interpret results:

```{r}
# Please ENTER the rotation eventually used (e.g. "none", "varimax", "quartimax", "promax", "oblimin", "simplimax", and "cluster" - see help(principal)). Defauls is "varimax"
selectizeInput("rotation_used", "",
               choices = c("none", "varimax", "quartimax", "bentlerT", "equamax",
                           "varimin", "geominT", "bifactor", "promax", "oblimin",
                           "simplimax", "bentlerQ", "geominQ", "biquartimin",
                           "cluster"),
               selected = "varimax")
```

For our data, the `r renderText(max(factors_selectedX()))` selected factors look as follows after the `r renderText(input$rotation_used)` rotation: 

```{r}
Rotated_FactorsX <- eventReactive({ 
  input$rotation_used
  factors_selectedX()
  ProjectDataFactorStandardized()
}, {
  Rotated_Results<-principal(ProjectDataFactorStandardized(), nfactors=max(factors_selectedX()), rotate=input$rotation_used,score=TRUE)
  Rotated_Factors<-round(Rotated_Results$loadings,2)
  Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
  colnames(Rotated_Factors)<-paste("Component",1:ncol(Rotated_Factors),sep=" ")

  sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
  Rotated_Factors <- Rotated_Factors[sorted_rows, , drop=F]
  Rotated_Factors
})

renderFormattable(dformat(Rotated_FactorsX()))
```

We can also "supress" loadings with small values. Please select the minimum values to display:

```{r}
# Please enter the minimum number below which you would like not to print - this makes the readability of the tables easier. Default values are either 10e6 (to print everything) or 0.5. Try both to see the difference.
sliderInput("MIN_VALUE", "", min=0, max=1, value=0.5)
```

Here are our rotated factors: 

```{r}
Rotated_Factors_thresX <- eventReactive({
  input$MIN_VALUE
  Rotated_FactorsX()
}, {
  Rotated_Factors_thres <- Rotated_FactorsX()
  Rotated_Factors_thres[abs(Rotated_Factors_thres) < input$MIN_VALUE]<-NA
  Rotated_Factors_thres
})

renderUI(tags$div(class="formattable_widget",HTML(gsub("NA", "", dformat(Rotated_Factors_thresX())))))
```


### Step 6: Save factor scores 

For our data, using the rotated factors we selected, we can create a new dataset where our observations are as follows (for the first `r SAMPLE_ROWS` observations):

```{r}
NEW_ProjectDataX <- eventReactive({
  input$rotation_used
  factors_selectedX()
  ProjectDataFactorStandardized()
},{
  Rotated_Results<-principal(ProjectDataFactorStandardized(), nfactors=max(factors_selectedX()), rotate=input$rotation_used,score=TRUE)
  NEW_ProjectData <- round(Rotated_Results$scores[,1:max(factors_selectedX()),drop=F],2)
  colnames(NEW_ProjectData)<-paste("Derived Variable (Factor)",1:ncol(NEW_ProjectData),sep=" ")
  NEW_ProjectData
})

renderFormattable(dformat(head(NEW_ProjectDataX(), SAMPLE_ROWS)))

renderUI(downloadButton("downloadData", "Save Dataset"))
output$downloadData <- downloadHandler( # TODO: use `outputArgs`
 filename = "dataset.csv",
 content = function(file) write.csv(NEW_ProjectDataX(), file)
)
```

We now can replace our original data with the new ones and continue our analysis. For example, we can now visualize our original data using only the newly derived attributes:

```{r}
tags$div(style='height:480px', renderC3({
  df <- as.data.frame(NEW_ProjectDataX())
  # TODO: An error occurs when ncol(df) > 9. Should investigate.
  df <- head(df[, head(colnames(df), 9), drop=F], SAMPLE_XPOINTS)
  colnames(df) <- paste("Derived Variable", 1:ncol(df))
  df <- cbind(
    list(observations = 1:nrow(df)),
    df
  )
  
  c3(melt(df, id="observations"), x="observations", y="value", group="variable") %>% c3_scatter()
}))
```

<div style="height: 450px;" class="tocify-extend-page" data-unique="tocify-extend-page">
</div>
