---

title : Data Analytics for Business
subtitle : Session 2-3, Dimensionality Reduction and Derived Attributes (Technical Slides)
author : T. Evgeniou 
job : INSEAD
widgets : []
mode : standalone 

---

## Why this class? Three goals
  
  
1. Organizational Level: Develop your intuition about identifying data analytics opportunities and their implementation challenges  
  
  
2. Project Level: Learn how to approach data analytics projects  
  
  
3. Analysis Level: Learn how to perform, read, and use key data analytics methods.


---

## Data Analytics Processes: Iterative Process Cycle with a Lifetime


  
Step 1: Defining the Objective(s)  
  
  
Step 2: Data Preparation and Exploration
  
  
Step 3: Analysis
  
  
Step 4: Performance Assessment



---

## What is Dimensionality Deduction and Factor Analysis?

Derive new  variables which are (linear) combinations of the original ones and capture most of the information in the original data. 

Is often used as a first step in Data Analytics

Can also be used to solve multicollinearity issues in regression

---

## Factor Analysis: Key idea


1. Transform the original selected variables into a smaller set of factors

2. Understand the underlying structure of the data and the new factors

3. Use the factors for subsequent analysis


---

## Key Questions

1. Can we really simplify the data by grouping the raw attributes?

2. How many factors should we use?

3. How good are the factors we found?

4. How interpretable and actionable are the factors we found? 

---

## Dimensionality Reduction and Factor Analysis: 6 (Easy) Steps

1. Confirm the data is metric 

2. Decide whether to scale or standardize the data

3. Check the correlation matrix to see if Factor Analysis makes sense

4. Develop a scree plot and decide on the number of factors to be derived

5. Interpret the factors (consider factor rotations - technical but useful)

6. Save factor scores for subsequent analyses

---

## Applying Factor Analysis: Evaluating MBA Applications

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# let's make the data into data.matrix classes so that we can easier visualize them
ProjectDataFactor = data.matrix(ProjectDataFactor)
ProjectData = data.matrix(ProjectData)
```

Variables available:

* GPA
* GMAT score
* Scholarships, fellowships won
* Evidence of Communications skills 
* Prior Job Experience
* Organizational Experience
* Other extra curricular achievements

Which variables are correlated? What do these variables capture?

---
## Example Factors

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
Rotated_Results<-principal(ProjectDataFactor, nfactors=max(factors_selected), rotate=rotation_used,score=TRUE)
Rotated_Factors<-round(Rotated_Results$loadings,2)
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Component",1:ncol(Rotated_Factors),sep=" ")

sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
Rotated_Factors <- Rotated_Factors[sorted_rows,]

show_data <- Rotated_Factors 
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---

## Step 1: Confirm the data are metric 


<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(ProjectDataFactor,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---

## Step 2: Decide whether to scale or standardize the data: Summary Statistics

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(my_summary(ProjectDataFactor),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>


---

## Data Standardization: Example Code

```{r, results='asis'}
ProjectDatafactor_scaled=apply(ProjectDataFactor,2, function(r) {
  if (sd(r)!=0) { 
    res=(r-mean(r))/sd(r) 
    } else { 
      res=0*r; res
      }
  })
```

---

## Standardized Data: Summary Statistics

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(my_summary(ProjectDatafactor_scaled),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---


## Step 3:  Check correlation matrix to see if Factor Analysis makes sense

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
thecor = round(cor(ProjectDataFactor),2)
colnames(thecor)<-colnames(thecor)
rownames(thecor)<-rownames(thecor)
## printing the result in a clean-slate table
#cat(renderHeatmapX(thecor, border=1,center = 0,vrange_up = 1, vrange_down = 1))
cat(renderHeatmapX(thecor, border=1))
```
</div>
</div>
</div>

---

## Step 4. Develop a scree plot and decide on the number of factors to be derived

For the method considered here (Principal Component Analysis):
* If there are n variables we will have n factors in total
* First factor will explain most of the variance, second next and so on.

Eigenvalues and Variance Explained by Factors
* each factor will have an associated eigenvalue -  which corresponds to the amount of variance explained by that factor
* with standardized variables each variable has a variance of 1, and the sum of all eigenvalues with n raw attributes is n
* we would like to capture as much of the total variance as possible, while keeping as few factors as possible

--- 

## How Many Factors? Eigenvalues and Variance Explained


<br>
<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, error=FALSE,message=FALSE,results='asis'}
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table

row=1:nrow(Variance_Explained_Table)
name<-paste("Component No:",row,sep="")
Variance_Explained_Table<-cbind(name,Variance_Explained_Table)
Variance_Explained_Table<-as.data.frame(Variance_Explained_Table)
colnames(Variance_Explained_Table)<-c("Components", "Eigenvalue", "Percentage_of_explained_variance", "Cumulative_percentage_of_explained_variance")

m<-gvisTable(Variance_Explained_Table,options=list(width=1200, height=400,allowHTML=TRUE,page='disable'),formats=list(Eigenvalue="#.##",Percentage_of_explained_variance="#.##",Cumulative_percentage_of_explained_variance="#.##"))
print(m,'chart')
```
</div>
</div>
</div>

--- 

## How Many Factors? Scree Plot

```{r Fig1, echo=FALSE, comment=NA, results='asis', message=FALSE, fig.align='center', fig=TRUE}

eigenvalues  <- Variance_Explained_Table[,2]
df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
colnames(df) <- c("eigenvalues", "components", "abline")
Line         <- gvisLineChart(as.data.frame(df), xvar="components", yvar=c("eigenvalues","abline"), options=list(title='Scree plot', legend="right", width=900, height=500, hAxis="{title:'Number of Components', titleTextStyle:{color:'black'}}", vAxes="[{title:'Eigenvalues'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(Line, 'chart')
```

```{r echo=FALSE, comment=NA, warning=FALSE,message=FALSE,results='asis'}
if (factor_selectionciterion == "eigenvalue")
  factors_selected = sum(Variance_Explained_Table_copy[,1] >= 1)
if (factor_selectionciterion == "variance")
  factors_selected = 1:head(which(Variance_Explained_Table_copy[,"cumulative percentage of variance"]>= minimum_variance_explained),1)
if (factor_selectionciterion == "manual")
  factors_selected = manual_numb_factors_used

```

--- 

## How many factors?

Three criteria to use:

* Eigenvalue > 1
* Cumulative variance explained
* "Elbow" in the Scree plot

Using the `r factor_selectionciterion` criterion we select `r factors_selected` factors. 

--- 

## Step 5. Interpret the factors 

Rotated Selected Factors using the `r rotation_used` rotation.

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
Rotated_Results<-principal(ProjectDataFactor, nfactors=max(factors_selected), rotate=rotation_used,score=TRUE)
Rotated_Factors<-round(Rotated_Results$loadings,2)
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Component",1:ncol(Rotated_Factors),sep=" ")

sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
Rotated_Factors <- Rotated_Factors[sorted_rows,]

show_data <- Rotated_Factors 
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---

## For visualization, let's supress the small numbers...

<br>

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
Rotated_Factors_thres <- Rotated_Factors
Rotated_Factors_thres[abs(Rotated_Factors_thres) < MIN_VALUE]<-NA
colnames(Rotated_Factors_thres)<- colnames(Rotated_Factors)
rownames(Rotated_Factors_thres)<- rownames(Rotated_Factors)

show_data <- Rotated_Factors_thres 
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---

## What Factor Loads "Look Good"? Three Technical Quality Criteria

1. For each factor (column) only a few loadings are large (in absolute value)

2. For each raw attribute (row) only a few loadings are large (in absolute value)

3. Any pair of factors (columns) should have different "patterns" of loading

---

## Step 6. Save factor scores for subsequent analyses

<br>

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
NEW_ProjectData <- round(Rotated_Results$scores[,1:factors_selected,drop=F],2)
colnames(NEW_ProjectData)<-paste("Derived Variable (Factor)",1:ncol(NEW_ProjectData),sep=" ")

if (factors_selected >=2){ 
  
  show_data <- as.data.frame(NEW_ProjectData) 
  show_data = show_data[1:min(max_data_report,nrow(show_data)),]
  row<-rownames(show_data)
  dfnew<-cbind(row,show_data)
  change<-colnames(dfnew)
  change[1]<-"Observation"
  colnames (dfnew)<-change
  m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
  print(m1,'chart')
  
  } else {
    print(xtable(NEW_ProjectData, caption="Only 1 derived variable generated", digits=3), type="html",html.table.attributes="class='table table-striped table-hover table-bordered'", caption.placement="top", comment=FALSE, include.rownames=TRUE, include.colnames=TRUE)
    }
```
</div>
</div>
</div>

---

## Using the Factor Scores: Perceptual Maps

```{r Fig2, echo=FALSE, comment=NA, results='asis', message=FALSE, echo=FALSE, fig.align='center', fig=TRUE}

if(ncol(NEW_ProjectData) >= 2) {
  df1  <- cbind(NEW_ProjectData[, 1], NEW_ProjectData[, 2])
  df1  <- as.data.frame(df1)
  sca1 <- gvisScatterChart(df1, options=list(legend="none",
                                             lineWidth=0, pointSize=8, hAxis.title="Derived Variable (Factor) 2",
                                             title="Data Visualization Using the top 2 Derived Attributes (Factors)", vAxis="{title:'Derived Variable (Factor) 1'}",
                                             hAxis="{title:'Derived Variable (Factor) 2'}", width=900, height=500))
  print(sca1,'chart')
  } else {
    df2  <- cbind(NEW_ProjectData[, 1], ProjectData[, 1])
    df2  <- as.data.frame(df2)
    sca2 <- gvisScatterChart(df2, options=list(legend="none",
                                               lineWidth=0, pointSize=12, hAxis.title="Derived Variable (Factor) 1",
                                               title="Only 1 Derived Variable: Using Initial Variable", vAxis="{title:'Derived Variable (Factor) 1'}",
                                               hAxis="{title:'Initial Variable (Factor) 1'}", width=900, height=800))
    print(sca2,'chart')
    }

```

---

## Factor Analysis: Some (Technical) Concepts 

1. Correlation
2. Variance explained (eigenvalues)
3. Scree plot
4. `r rotation_used` rotation
5. Factor Loadings ("components")
6. Factor scores

---

## Key Questions

1. How many factors should we use? Why? Quantitative and Qualitative criteria

2. How can we name and interpret the factors? 

3. What are some issues to consider? 

---

## Next Class: Cluster Analysis for Segmentation

```{r include=FALSE}
## using dummy data to recreate cluster segmentation graphics
library(vegan)
require(vegan)
data(dune)
# kmeans
kclus <- kmeans(dune,centers= 4, iter.max=1000)
# distance matrix
dune_dist <- dist(dune,method="euclidian")
# Multidimensional scaling
cmd <- cmdscale(dune_dist)
```

```{r fig.width=12, fig.height=6, message=FALSE, echo=FALSE, fig.align='center', warning=FALSE, fig=TRUE}
# plot MDS, with colors by groups from kmeans
groups <- levels(factor(kclus$cluster))
ordiplot(cmd, type = "n")
cols <- c("steelblue", "darkred", "darkgreen", "pink")
for(i in seq_along(groups)){
  points(cmd[factor(kclus$cluster) == groups[i], ], col = cols[i], pch = 16)
  }

# add spider and hull
ordispider(cmd, factor(kclus$cluster), label = TRUE)
ordihull(cmd, factor(kclus$cluster), lty = "dotted")
```
