<link rel="stylesheet" href="http://netdna.bootstrapcdn.com/bootstrap/3.0.3/css/bootstrap.min.css">
<style type="text/css"> body {padding: 10px 30px 10px 30px;} table,th, td {text-align: center;}</style>
<style>
td.tableRow
{
text-align:center;
}
</style>
<script src='//api.tiles.mapbox.com/mapbox.js/v1.6.0/mapbox.js'></script>


Boats Segmentation: Sample Segments
========================================================

**T. Evgeniou**

No Segmentation Specific Solution
-----------------------------------------------------------------------------

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# let's make the data into data.matrix classes so that we can easier visualize them
ProjectData = data.matrix(ProjectData)

if (random_sampling){
  estimation_data_ids=sample.int(nrow(ProjectData),floor(estimation_data_percent*nrow(ProjectData)/100))
  non_estimation_data = setdiff(1:nrow(ProjectData),estimation_data_ids)
  validation_data_ids=non_estimation_data[sample.int(length(non_estimation_data), floor(validation_data_percent/(validation_data_percent+test_data_percent)*length(non_estimation_data)))]
  } else {
    estimation_data_ids=1:floor(estimation_data_percent*nrow(ProjectData)/100)
    non_estimation_data = setdiff(1:nrow(ProjectData),estimation_data_ids)
    validation_data_ids = (tail(estimation_data_ids,1)+1):(tail(estimation_data_ids,1) + floor(validation_data_percent/(validation_data_percent+test_data_percent)*length(non_estimation_data)))
    }

test_data_ids = setdiff(1:nrow(ProjectData), union(estimation_data_ids,validation_data_ids))

estimation_data=ProjectData[estimation_data_ids,]
validation_data=ProjectData[validation_data_ids,]
test_data=ProjectData[test_data_ids,]

# just name the variables numerically so that they look ok on the tree plots
independent_variables_nolabel = paste("IV", 1:length(independent_variables), sep="")

estimation_data_nolabel = cbind(estimation_data[,dependent_variable], estimation_data[,independent_variables])
colnames(estimation_data_nolabel)<- c(colnames(estimation_data)[dependent_variable],independent_variables_nolabel)

validation_data_nolabel = cbind(validation_data[,dependent_variable], validation_data[,independent_variables])
colnames(validation_data_nolabel)<- c(dependent_variable,independent_variables_nolabel)

test_data_nolabel = cbind(test_data[,dependent_variable], test_data[,independent_variables])
colnames(test_data_nolabel)<- c(dependent_variable,independent_variables_nolabel)

estimation_data_nolabel = data.frame(estimation_data_nolabel)
validation_data_nolabel = data.frame(validation_data_nolabel)
test_data_nolabel = data.frame(test_data_nolabel)

estimation_data = data.frame(estimation_data)
validation_data = data.frame(validation_data)
test_data = data.frame(test_data)

formula=paste(colnames(estimation_data)[dependent_variable],paste(Reduce(paste,sapply(head(independent_variables_nolabel,-1), function(i) paste(i,"+",sep=""))),tail(independent_variables_nolabel,1),sep=""),sep="~")
CART_tree<-rpart(formula, data= estimation_data_nolabel,method="class", control=CART_control)

fancyRpartPlot(CART_tree)
```
</center>


<center style="width=900px; height=800px;">
```{r echo=FALSE, message=FALSE,warning=FALSE, prompt=FALSE, results='asis',fig.width=14,fig.height=10}
CART_tree_large<-rpart(formula, data= estimation_data_nolabel,method="class", control=rpart.control(cp = 0.005))

fancyRpartPlot(CART_tree_large)
```
</center>

```{r echo=FALSE, message=FALSE,warning=FALSE, results='asis'}
# Let's first calculate all probabilites for the estimation, validation, and test data
estimation_Probability_class1_tree<-predict(CART_tree, estimation_data_nolabel)[,2]
estimation_Probability_class1_tree_large<-predict(CART_tree_large, estimation_data_nolabel)[,2]

validation_Probability_class1_tree<-predict(CART_tree, validation_data_nolabel)[,2]
validation_Probability_class1_tree_large<-predict(CART_tree_large, validation_data_nolabel)[,2]

test_Probability_class1_tree<-predict(CART_tree, test_data_nolabel)[,2]
test_Probability_class1_tree_large<-predict(CART_tree_large, test_data_nolabel)[,2]


estimation_prediction_class_tree=1*as.vector(estimation_Probability_class1_tree > Probability_Threshold)
estimation_prediction_class_tree_large=1*as.vector(estimation_Probability_class1_tree_large > Probability_Threshold)

validation_prediction_class_tree=1*as.vector(validation_Probability_class1_tree > Probability_Threshold)
validation_prediction_class_tree_large=1*as.vector(validation_Probability_class1_tree_large > Probability_Threshold)

test_prediction_class_tree=1*as.vector(test_Probability_class1_tree > Probability_Threshold)
test_prediction_class_tree_large=1*as.vector(test_Probability_class1_tree_large > Probability_Threshold)


Classification_Table=rbind(validation_data[,dependent_variable],validation_Probability_class1_tree)
rownames(Classification_Table)<-c("Actual Class","Probability of Class 1")
colnames(Classification_Table)<- paste("Obs", 1:ncol(Classification_Table), sep=" ")

Classification_Table_large=rbind(validation_data[,dependent_variable],validation_Probability_class1_tree)
rownames(Classification_Table_large)<-c("Actual Class","Probability of Class 1")
colnames(Classification_Table_large)<- paste("Obs", 1:ncol(Classification_Table_large), sep=" ")


formula_log=paste(colnames(estimation_data[,dependent_variable,drop=F]),paste(Reduce(paste,sapply(head(independent_variables,-1), function(i) paste(colnames(estimation_data)[i],"+",sep=""))),colnames(estimation_data)[tail(independent_variables,1)],sep=""),sep="~")

logreg_solution <- glm(formula_log, family=binomial(link="logit"),  data=estimation_data)

log_coefficients = round(summary(logreg_solution)$coefficients,1)

```

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# Let's get the probabilities for the 3 types of data again
estimation_Probability_class1_log<-predict(logreg_solution, type="response", newdata=estimation_data[,independent_variables])
validation_Probability_class1_log<-predict(logreg_solution, type="response", newdata=validation_data[,independent_variables])
test_Probability_class1_log<-predict(logreg_solution, type="response", newdata=test_data[,independent_variables])

estimation_prediction_class_log=1*as.vector(estimation_Probability_class1_log > Probability_Threshold)
validation_prediction_class_log=1*as.vector(validation_Probability_class1_log > Probability_Threshold)
test_prediction_class_log=1*as.vector(test_Probability_class1_log > Probability_Threshold)

Classification_Table=rbind(validation_data[,dependent_variable],validation_Probability_class1_log)
rownames(Classification_Table)<-c("Actual Class","Probability of Class 1")
colnames(Classification_Table)<- paste("Obs", 1:ncol(Classification_Table), sep=" ")

log_importance = tail(log_coefficients[,"z value", drop=F],-1) # remove the intercept
log_importance = log_importance/max(abs(log_importance))

tree_importance = CART_tree$variable.importance
tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree$variable.importance)))
tree_importance_final = rep(0,length(independent_variables))
tree_importance_final[tree_ordered_drivers] <- tree_importance
tree_importance_final <- tree_importance_final/max(abs(tree_importance_final))
tree_importance_final <- tree_importance_final*sign(log_importance)

large_tree_importance = CART_tree_large$variable.importance
large_tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree_large$variable.importance)))
large_tree_importance_final = rep(0,length(independent_variables))
large_tree_importance_final[large_tree_ordered_drivers] <- large_tree_importance
large_tree_importance_final <- large_tree_importance_final/max(abs(large_tree_importance_final))
large_tree_importance_final <- large_tree_importance_final*sign(log_importance)

Importance_table <- cbind(tree_importance_final,large_tree_importance_final, log_importance)
colnames(Importance_table) <- c("CART 1", "CART 2", "Logistic Regr.")
rownames(Importance_table) <- rownames(log_importance)
## printing the result in a clean-slate table
cat(renderHeatmapX(Importance_table, border=1, center = 0, minvalue = 0))
```

The profit curves for the test data in this case are:

<style>
.wrapper{


width: 100%;

overflow-x: scroll;

}
.wrapper1{

height:450px;
overflow-y: scroll;
}
</style>
<div class="wrapper wrapper1">

```{r echo=FALSE,results='asis',warning=FALSE,error=FALSE}

actual_class<- test_data[,dependent_variable]

probs = test_Probability_class1_tree
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev1   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data CART 1', legend="right", width=600, height=600, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev1,'chart')

probs = test_Probability_class1_tree_large
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev2   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data CART 2', legend="right", width=600, height=400, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev2,'chart')

probs = test_Probability_class1_log
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev3   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data logistic regression', legend="right", width=600, height=400, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev3,'chart')
```
</div>

<br>
<br>
<hr>

Segment Specific Analysis: Five Segments Solution
---------------------------------------------------------------------------

Let's see first how many observations we have in each segment:

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
numb_clusters_used = 5

cluster_file = paste(paste(local_directory,"data", sep="/"),paste(paste(datafile_name,paste("cluster",numb_clusters_used,sep="") , sep="_"), "csv", sep="."), sep="/")

cluster_ids <- read.csv(cluster_file, sep=",", dec=".") # this contains only the matrix ProjectData
cluster_ids <- data.matrix(cluster_ids)
cluster_ids = cluster_ids[,2]

cluster_size = NULL
for (i in sort(unique(cluster_ids))){
  cluster_size = c(cluster_size,sum(cluster_ids == i))
  }
cluster_size = matrix(cluster_size, nrow=1)
colnames(cluster_size) <- paste("Segment", 1:length(cluster_size), sep=" ")
rownames(cluster_size) <- "Number of Obs."
print(xtable(cluster_size ,caption="Number of Observations per segment in the Dataset", digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)
```
</div>
</div>

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
actual_class<- test_data[,dependent_variable]
probs_tree = 0*test_Probability_class1_tree
probs_tree_large = 0*test_Probability_class1_tree_large
probs_log = 0*test_Probability_class1_log
Log_Drivers = NULL

for (i in sort(unique(cluster_ids))){
  useonly = which(cluster_ids==i)
  if (length(useonly) >= min_segment){
    
    test_ids_used = intersect(test_data_ids,useonly)
    probs_to_fill = which(sapply(test_data_ids, function(i) sum(test_ids_used==i)) !=0)
    estimation_data_clus=ProjectData[intersect(estimation_data_ids,useonly) ,]
    test_data_clus=ProjectData[intersect(test_data_ids,useonly),]
    
    ###
    estimation_data_clus_nolabel = cbind(estimation_data_clus[,dependent_variable], estimation_data_clus[,independent_variables])
    colnames(estimation_data_clus_nolabel)<- c(colnames(estimation_data_clus)[dependent_variable],independent_variables_nolabel)
    
    test_data_clus_nolabel = cbind(test_data_clus[,dependent_variable], test_data_clus[,independent_variables])
    colnames(test_data_clus_nolabel)<- c(dependent_variable,independent_variables_nolabel)
    
    estimation_data_clus = data.frame(estimation_data_clus)
    test_data_clus = data.frame(test_data_clus)
    estimation_data_clus_nolabel = data.frame(estimation_data_clus_nolabel)
    test_data_clus_nolabel = data.frame(test_data_clus_nolabel)
    ###
    
    CART_tree<-rpart(formula, data= estimation_data_clus_nolabel,method="class", control=CART_control)    
    CART_tree_large<-rpart(formula, data= estimation_data_clus_nolabel,method="class", control=rpart.control(cp = 0.005))
    logreg_solution <- glm(formula_log, family=binomial(link="logit"),  data=estimation_data_clus)
    
    #####
    
    test_Probability_class1_tree<-predict(CART_tree, test_data_clus_nolabel)[,2]
    test_Probability_class1_tree_large<-predict(CART_tree_large, test_data_clus_nolabel)[,2]
    test_Probability_class1_log<-predict(logreg_solution, type="response", newdata=test_data_clus[,independent_variables])
    
    #######
    probs_tree[probs_to_fill] <- test_Probability_class1_tree
    probs_tree_large[probs_to_fill] <- test_Probability_class1_tree
    probs_log[probs_to_fill] <- test_Probability_class1_log
    
    
    log_coefficients = round(summary(logreg_solution)$coefficients,1)
    Log_Drivers_segment = tail(log_coefficients[,"z value", drop=F],-1) # remove the intercept
    Log_Drivers_segment = Log_Drivers_segment/max(abs(Log_Drivers_segment))
    
    tree_importance = CART_tree$variable.importance
    tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree$variable.importance)))
    tree_importance_final = rep(0,length(independent_variables))
    tree_importance_final[tree_ordered_drivers] <- tree_importance
    tree_importance_final <- tree_importance_final/max(abs(tree_importance_final))
    tree_importance_final <- tree_importance_final*sign(Log_Drivers_segment)
    
    #Log_Drivers = cbind(Log_Drivers,tree_importance_final)
    Log_Drivers = cbind(Log_Drivers,Log_Drivers_segment)

    
    }
  }
colnames(Log_Drivers) <- paste("Segment", 1:length(unique(cluster_ids)), sep = " ")
cat(renderHeatmapX(tail(Log_Drivers,-1), border=1, center = 0, minvalue = 0))

```
<br>
<br>

The profit curves for the test data in this case are:

<style>
.wrapper{


width: 100%;

overflow-x: scroll;

}
.wrapper1{

height:450px;
overflow-y: scroll;
}
</style>
<div class="wrapper wrapper1">

```{r echo=FALSE,results='asis',warning=FALSE,error=FALSE}

actual_class<- test_data[,dependent_variable]

probs = probs_tree
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev1   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data CART 1', legend="right", width=600, height=600, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev1,'chart')

probs = probs_tree_large
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev2   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data CART 2', legend="right", width=600, height=400, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev2,'chart')

probs = probs_log
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev3   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data logistic regression', legend="right", width=600, height=400, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev3,'chart')
```
</div>

<br>
<br>
<br>
<hr>


Segment Specific Analysis: Seven Segments Solution
---------------------------------------------------------------------------

Let's see first how many observations we have in each segment:

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
numb_clusters_used = 7

cluster_file = paste(paste(local_directory,"data", sep="/"),paste(paste(datafile_name,paste("cluster",numb_clusters_used,sep="") , sep="_"), "csv", sep="."), sep="/")

cluster_ids <- read.csv(cluster_file, sep=",", dec=".") # this contains only the matrix ProjectData
cluster_ids <- data.matrix(cluster_ids)
cluster_ids = cluster_ids[,2]

cluster_size = NULL
for (i in sort(unique(cluster_ids))){
  cluster_size = c(cluster_size,sum(cluster_ids == i))
  }
cluster_size = matrix(cluster_size, nrow=1)
colnames(cluster_size) <- paste("Segment", 1:length(cluster_size), sep=" ")
rownames(cluster_size) <- "Number of Obs."
print(xtable(cluster_size ,caption="Number of Observations per segment in the Dataset", digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)
```
</div>
</div>

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
actual_class<- test_data[,dependent_variable]
probs_tree = 0*test_Probability_class1_tree
probs_tree_large = 0*test_Probability_class1_tree_large
probs_log = 0*test_Probability_class1_log
Log_Drivers = NULL

for (i in sort(unique(cluster_ids))){
  useonly = which(cluster_ids==i)
  if (length(useonly) >= min_segment){
    
    test_ids_used = intersect(test_data_ids,useonly)
    probs_to_fill = which(sapply(test_data_ids, function(i) sum(test_ids_used==i)) !=0)
    estimation_data_clus=ProjectData[intersect(estimation_data_ids,useonly) ,]
    test_data_clus=ProjectData[intersect(test_data_ids,useonly),]
    
    ###
    estimation_data_clus_nolabel = cbind(estimation_data_clus[,dependent_variable], estimation_data_clus[,independent_variables])
    colnames(estimation_data_clus_nolabel)<- c(colnames(estimation_data_clus)[dependent_variable],independent_variables_nolabel)
    
    test_data_clus_nolabel = cbind(test_data_clus[,dependent_variable], test_data_clus[,independent_variables])
    colnames(test_data_clus_nolabel)<- c(dependent_variable,independent_variables_nolabel)
    
    estimation_data_clus = data.frame(estimation_data_clus)
    test_data_clus = data.frame(test_data_clus)
    estimation_data_clus_nolabel = data.frame(estimation_data_clus_nolabel)
    test_data_clus_nolabel = data.frame(test_data_clus_nolabel)
    ###
    
    CART_tree<-rpart(formula, data= estimation_data_clus_nolabel,method="class", control=CART_control)    
    CART_tree_large<-rpart(formula, data= estimation_data_clus_nolabel,method="class", control=rpart.control(cp = 0.005))
    logreg_solution <- glm(formula_log, family=binomial(link="logit"),  data=estimation_data_clus)
    
    #####
    
    test_Probability_class1_tree<-predict(CART_tree, test_data_clus_nolabel)[,2]
    test_Probability_class1_tree_large<-predict(CART_tree_large, test_data_clus_nolabel)[,2]
    test_Probability_class1_log<-predict(logreg_solution, type="response", newdata=test_data_clus[,independent_variables])
    
    #######
    probs_tree[probs_to_fill] <- test_Probability_class1_tree
    probs_tree_large[probs_to_fill] <- test_Probability_class1_tree
    probs_log[probs_to_fill] <- test_Probability_class1_log
    
    
    log_coefficients = round(summary(logreg_solution)$coefficients,1)
    Log_Drivers_segment = tail(log_coefficients[,"z value", drop=F],-1) # remove the intercept
    Log_Drivers_segment = Log_Drivers_segment/max(abs(Log_Drivers_segment))
    
    tree_importance = CART_tree$variable.importance
    tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree$variable.importance)))
    tree_importance_final = rep(0,length(independent_variables))
    tree_importance_final[tree_ordered_drivers] <- tree_importance
    tree_importance_final <- tree_importance_final/max(abs(tree_importance_final))
    tree_importance_final <- tree_importance_final*sign(Log_Drivers_segment)
    
    Log_Drivers = cbind(Log_Drivers,tree_importance_final)

    
    }
  }
colnames(Log_Drivers) <- paste("Segment", 1:length(unique(cluster_ids)), sep = " ")
cat(renderHeatmapX(tail(Log_Drivers,-1), border=1, center = 0, minvalue = 0))

```
<br>
<br>

The profit curves in this case are:

<style>
.wrapper{


width: 100%;

overflow-x: scroll;

}
.wrapper1{

height:450px;
overflow-y: scroll;
}
</style>
<div class="wrapper wrapper1">

```{r echo=FALSE,results='asis',warning=FALSE,error=FALSE}

actual_class<- test_data[,dependent_variable]

probs = probs_tree
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev1   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data CART 1', legend="right", width=600, height=600, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev1,'chart')

probs = probs_tree_large
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev2   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data CART 2', legend="right", width=600, height=400, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev2,'chart')

probs = probs_log
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
frame<-cbind(xaxis,yaxis)
frame<-as.data.frame(frame)
LineFramev3   <- gvisLineChart(frame, xvar=c("xaxis"), yvar="yaxis", options=list(title='Profit Curve for test data logistic regression', legend="right", width=600, height=400, hAxis="{title:'Percent Selected', titleTextStyle:{color:'black'}}", vAxes="[{title:'Estimated Profit'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(LineFramev3,'chart')
```
</div>

<br>
<br>
