---

title : Data Analytics for Business
subtitle : Sessions 4-5, Clustering and Segmentation
author : T. Evgeniou 
job : INSEAD
widgets : []
mode : standalone 

---

## Clustering and Segmentation

```{r include=FALSE}
## using dummy data to recreate cluster segmentation graphics
library(vegan)
require(vegan)
data(dune)
# kmeans
kclus <- kmeans(dune,centers= 4, iter.max=1000)
# distance matrix
dune_dist <- dist(dune,method=distance_used)
# Multidimensional scaling
cmd <- cmdscale(dune_dist)
```

```{r fig.width=12, fig.height=6, message=FALSE, echo=FALSE, fig.align='center', warning=FALSE, fig=TRUE}
# plot MDS, with colors by groups from kmeans
groups <- levels(factor(kclus$cluster))
ordiplot(cmd, type = "n")
cols <- c("steelblue", "darkred", "darkgreen", "pink")
for(i in seq_along(groups)){
  points(cmd[factor(kclus$cluster) == groups[i], ], col = cols[i], pch = 16)
  }

# add spider and hull
ordispider(cmd, factor(kclus$cluster), label = TRUE)
ordihull(cmd, factor(kclus$cluster), lty = "dotted")
```


---

## What is Clustering and Segmentation?

<br>
<br>
<br>

Processes and tools to organize data in a few segments, with data being as similar as possible within each segment, and as different as possible across segments

---

## Example Usage

<br>
<br>

- Market Segmentation

- Co-Moving Asset Classes

- Geo-demographic segmentation

- Recommender Systems

- Text Mining


---

## A Segmentation Process

<br>

1. Confirm the data in metric 

2. Decide whether to scale or standardize the data

3. Decide which variables to use for clustering

4. Define similarity or dissimilarity measures between observations

5. Visualize Individual Attributes and  Pair-wise Distances between the Observations

6. Select the clustering method to use and decide how many clusters to have

7. Profile and interpret the clusters 

8. Assess the robustness of our clusters

---

## Example Data: Market Research Survey


```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# let's make the data into data.matrix classes so that we can easier visualize them
ProjectData_segment = data.matrix(ProjectData_segment)
ProjectData_profile = data.matrix(ProjectData_profile)
ProjectData = data.matrix(ProjectData)
```

Size of data: `r nrow(ProjectData)` observations with `r ncol(ProjectData)` variables each

V1: Shopping is fun (scale 1-7)

V2: Shopping is bad for your budget (scale 1-7)

V3: I combine shopping with eating out (scale 1-7)

V4: I try to get the best buys while shopping (scale 1-7)

V5: I don't care about shopping (scale 1-7)

V6: You can save lot of money by comparingprices (scale 1-7)

Income: the household income of the respondent (in dollars)

Mall.Visits: how often they visit the mall (scale 1-7)

---

## Step 1: Confirm the data are metric 


<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(ProjectData_segment,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---

## Step 2: Decide whether to scale or standardize the data: Summary Statistics

<br>

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(my_summary(ProjectData_segment),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>


---

## Data Standardization: Example Code

<br>
<br>

```{r, results='asis'}
ProjectData_segment_scaled=apply(ProjectData_segment,2, function(r) {
  if (sd(r)!=0) { 
    res=(r-mean(r))/sd(r) 
    } else { 
      res=0*r; res
      }
  })
```

---

## Standardized Data: Summary Statistics

<br>

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(my_summary(ProjectData_segment_scaled),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---

## Step 3. Decide which variables to use for clustering

<br>
<br>
<br>

The choice of the variables used for clustering is critically important 
<br>
<br>

Tpically we use different variables for segmentation (the "segmentation variables") and different ones for profiling (the "profiling variables")

<br>
<br>

Remember: Segmentation is an iterative process

---

## Step 4. Define similarity or dissimilarity measures between observations


Defining what we mean when we say "similar" or "different" observations is a key part of cluster analysis which often requires a lot of contextual knowledge and creativity 

<br>
<br>


There are literally thousands of rigorous mathematical definitions of distance between observations/vectors

<br>
<br>

The user can manually define such distance metrics

---

## Distances across our data using the Euclidean distance

<br>
<br>

```{r include=FALSE, echo=FALSE, comment=NA, warning=FALSE, message=FALSE}
euclidean_pairwise <- as.matrix(dist(head(ProjectData_segment, 5), method="euclidean"))
euclidean_pairwise <- euclidean_pairwise*lower.tri(euclidean_pairwise) + euclidean_pairwise*diag(euclidean_pairwise) + 10e10*upper.tri(euclidean_pairwise)
euclidean_pairwise[euclidean_pairwise==10e10] <- NA
```

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
print(xtable(euclidean_pairwise, caption="Pairwise Distances between the first 5 observations using The Euclidean Distance Metric", digits=1), type="html", html.table.attributes = "class='table table-striped table-hover table-bordered'", caption.placement="top", comment = FALSE, include.rownames = FALSE)
```
</div>
</div>
</div>


---

## Distances across our data using the Manhattan distance

<br>
<br>

```{r include=FALSE, echo=FALSE, comment=NA, warning=FALSE, message=FALSE}
manhattan_pairwise <- as.matrix(dist(head(ProjectData_segment, 5), method="manhattan"))
manhattan_pairwise <- manhattan_pairwise*lower.tri(manhattan_pairwise) + manhattan_pairwise*diag(manhattan_pairwise) + 10e10*upper.tri(manhattan_pairwise)
manhattan_pairwise[manhattan_pairwise==10e10] <- NA
```

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
print(xtable(manhattan_pairwise, caption="Pairwise Distances between the first 5 observations using The Manhattan Distance Metric", digits=1), type="html", html.table.attributes = "class='table table-striped table-hover table-bordered'", caption.placement="top", comment = FALSE, include.rownames = FALSE)
```
</div>
</div>
</div>


---


## Manually Defined Distances: an Example

<br>
<br>
<br>

```{r ,results='asis'}
My_Distance_function<-function(x,y){sum(abs(x-y)>2)}
```

---


## Manually Defined Distances: an Example

<br>
<br>

```{r include=FALSE, echo=FALSE, comment=NA, warning=FALSE, fig.align='center', message=FALSE}
Manual_Pairwise=apply(head(ProjectData_segment,5),1,function(i) apply(head(ProjectData_segment,5),1,function(j) My_Distance_function(i,j) ))
Manual_Pairwise <- Manual_Pairwise * lower.tri(Manual_Pairwise) + Manual_Pairwise * diag(Manual_Pairwise) + 10e10*upper.tri(Manual_Pairwise)
Manual_Pairwise[Manual_Pairwise == 10e10] <- NA
```

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
print(xtable(Manual_Pairwise, caption="Pairwise Distances between the first 5 observations using a simple manually defined Distance Metric", digits=1), type="html", html.table.attributes = "class='table table-striped table-hover table-bordered'", caption.placement="top", comment = FALSE, include.rownames = FALSE)
```
</div>
</div>
</div>


---

## Step 5. Visualize Individual Attributes and  Pair-wise Distances between the Observations

---


## Histogram of all pairwise distances 

Distance Used:  `r distance_used` 
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
Pairwise_Distances <- dist(ProjectData_segment, method = distance_used) 
hist(Pairwise_Distances, main = NULL, xlab="Histogram of all pairwise Distances between observtions", ylab="Frequency")
```


---

## Step 6. Select the clustering method to use and decide how many clusters to have

<br>
<br>

There are many clustering methods. Two common ones are:

* Hierarchical Methods

* Non-Hierarchical Methods (e.g. k-means)

<br>

We can plug-and-play (with CARE) this "black box" in our analysis 

---


## Hierarchical Clustering: Dendrogram

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
Hierarchical_Cluster_distances <- dist(ProjectData_segment, method=distance_used)
Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=hclust_method)
# Display dendogram
plot(Hierarchical_Cluster, main = NULL, sub=NULL, labels = 1:nrow(ProjectData_segment), xlab="Our Observations", cex.lab=1, cex.axis=1) 
# Draw dendogram with red borders around the 3 clusters
rect.hclust(Hierarchical_Cluster, k=numb_clusters_used, border="red") 
```
<div>

---

## Hierarchical Clustering Dendrogram Heights Plot

<style>
.wrapper{
height: 40%;
width: 800px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:15px;">
```{r Fig1, echo=FALSE, comment=NA, results='asis', message=FALSE, fig.align='center', fig=TRUE}
max <- nrow(ProjectData)
num <- max - 1
df1 <- cbind(as.data.frame(Hierarchical_Cluster$height[length(Hierarchical_Cluster$height):1]), c(1:num))
colnames(df1) <- c("distances","index")
Line <- gvisLineChart(as.data.frame(df1), xvar="index", yvar="distances", options=list(title='Distances plot', legend="right", width=900, height=600, hAxis="{title:'Number of Components', titleTextStyle:{color:'black'}}", vAxes="[{title:'Distances'}]", series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(Line,'chart')
```
<div>

---

## Cluster Membership: Hierarchical Clustering

<br>

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
cluster_memberships_hclust <- as.vector(cutree(Hierarchical_Cluster, k=numb_clusters_used)) # cut tree into 3 clusters
cluster_ids_hclust=unique(cluster_memberships_hclust)
ProjectData_with_hclust_membership <- cbind(1:length(cluster_memberships_hclust),cluster_memberships_hclust)
colnames(ProjectData_with_hclust_membership)<-c("Observation Number","Cluster_Membership")
```


<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectData_with_hclust_membership,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Observation"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---

## Cluster Membership: K-means Clustering

Note: K-means does not necessarily lead to the same solution every time you run it

<br>

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
kmeans_clusters <- kmeans(ProjectData_segment,centers= numb_clusters_used, iter.max=2000, algorithm=kmeans_method)

ProjectData_with_kmeans_membership <- cbind(1:length(kmeans_clusters$cluster),kmeans_clusters$cluster)
colnames(ProjectData_with_kmeans_membership)<-c("Observation Number","Cluster_Membership")
```


<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectData_with_kmeans_membership,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Observation"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
</div>

---

## Step 7. Profile and interpret the clusters 

<br>
<br>
<br>

Data analytics is used to eventually make decisions, and that is feasible only when we are comfortable (enough) with our understanding of the analytics results, including our ability to clearly interpret them. 

---

## Cluster Profiling: Cluster Centers of Profling Variables


<br>

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
cluster_memberships_kmeans <- kmeans_clusters$cluster 
cluster_ids_kmeans <- unique(cluster_memberships_kmeans)
```

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}

cluster_memberships <- cluster_memberships_hclust
cluster_ids <-  cluster_ids_hclust  
if (profile_with == "hclust"){
  cluster_memberships <- cluster_memberships_hclust
  cluster_ids <-  cluster_ids_hclust  
  }
if (profile_with == "kmeans"){
  cluster_memberships <- cluster_memberships_kmeans
  cluster_ids <-  cluster_ids_kmeans
  }

population_average = matrix(apply(ProjectData_profile, 2, mean), ncol=1)
colnames(population_average) <- "Population"
Cluster_Profile_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_profile[(cluster_memberships==i), ], 2, mean))
if (ncol(ProjectData_profile) <2)
  Cluster_Profile_mean=t(Cluster_Profile_mean)
colnames(Cluster_Profile_mean) <- paste("Segment", 1:length(cluster_ids), sep=" ")
cluster.profile <- cbind (population_average,Cluster_Profile_mean)
```

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-6">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
show_data = data.frame(round(cluster.profile,2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')

```
</div>
</div>
</div>

--- 

## Interpretation: Snake Plots

<style>
.wrapper{
height: 160%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
```{r Fig2, fig.width=6, fig.height=6, message=FALSE, echo=FALSE, fig.align='center', warning=FALSE, fig=TRUE}
ProjectData_scaled=apply(ProjectData,2, function(r) {if (sd(r)!=0) res=(r-mean(r))/sd(r) else res=0*r; res})
ProjectData_scaled_profile = ProjectData_scaled[, profile_attributes_used,drop=F]

Cluster_Profile_standar_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_scaled_profile[(cluster_memberships==i), ,drop = F], 2, mean))
if (ncol(ProjectData_scaled_profile) < 2)
  Cluster_Profile_standar_mean = t(Cluster_Profile_standar_mean)
colnames(Cluster_Profile_standar_mean) <- paste("Segment", 1:length(cluster_ids), sep=" ")

plot(Cluster_Profile_standar_mean[, 1,drop=F], type="l", col="red", main="Snake plot for each cluster", ylab="mean of cluster", xlab="profiling variables (standardized)",ylim=c(min(Cluster_Profile_standar_mean),max(Cluster_Profile_standar_mean))) 
for(i in 2:ncol(Cluster_Profile_standar_mean))
  lines(Cluster_Profile_standar_mean[, i], col="blue")
```
</div>


--- 

## Interpretation: Ratio to Average of Total Population (1 = Average)


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
population_average_matrix <- population_average[,"Population",drop=F] %*% matrix(rep(1,ncol(Cluster_Profile_mean)),nrow=1)
cluster_profile_ratios <- (ifelse(population_average_matrix==0, 0,Cluster_Profile_mean/population_average_matrix))
colnames(cluster_profile_ratios) <- paste("Segment", 1:ncol(cluster_profile_ratios), sep=" ")
rownames(cluster_profile_ratios) <- colnames(ProjectData)[profile_attributes_used]
## printing the result in a clean-slate table
```

<style>
.wrapper{
height: 120%;
width: 900px;
overflow: auto;
}
</style>
<div class="wrapper" style="font-size:20px;">
<div class="row">
<div class="col-md-6">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
show_data = data.frame(round(cluster_profile_ratios,2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
#print(m1,'chart')
cat(renderHeatmapX(cluster_profile_ratios, border=1, center = 1, minvalue = heatmin))
```
</div>
</div>
</div>

---

## The data

Size of data: `r nrow(ProjectData)` observations with `r ncol(ProjectData)` variables each

V1: Shopping is fun (scale 1-7)

V2: Shopping is bad for your budget (scale 1-7)

V3: I combine shopping with eating out (scale 1-7)

V4: I try to get the best buys while shopping (scale 1-7)

V5: I don't care about shopping (scale 1-7)

V6: You can save lot of money by comparingprices (scale 1-7)

Income: the household income of the respondent (in dollars)

Mall.Visits: how often they visit the mall (scale 1-7)

--- 

## Step 8. Assess the robustness of our clusters

<br>
<br>

Are the clusters stable when we use:
<br>
<br>

- using different subsets of the original data
- using variations of the original segmentation attributes
- using different distance metrics
- using different segmentation methods
- using different numbers of clusters


--- 

## Example Robustness Test: Different Methods

<br>
<br>

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
# First, make sure the segment ids are correctly aligned
cluster_overlaps <- Reduce(cbind,lapply(1:length(cluster_ids_kmeans), function(i) {
  overlaps <- sapply(1:length(cluster_ids_hclust), function(j) {
    length(intersect(which(cluster_memberships_kmeans==i), 
                     which(cluster_memberships_hclust==j))) } );
  overlaps}))
max_cluster_overlap = rep(0,length(cluster_ids_kmeans))
for (i in 1:length(cluster_ids_kmeans)){
  highest_now = which.max(cluster_overlaps)
  hclust_id_now = highest_now %% length(cluster_ids_kmeans)
  hclust_id_now = ifelse(hclust_id_now == 0, 3, hclust_id_now)
  kmeans_id_now = ceiling(highest_now/length(cluster_ids_kmeans))
  max_cluster_overlap[kmeans_id_now] <- hclust_id_now
  cluster_overlaps[hclust_id_now,] <- 0
  cluster_overlaps[,kmeans_id_now] <- 0
}
cluster_memberships_kmeans_aligned <- rep(0,length(cluster_memberships_kmeans))
for (i in 1:length(cluster_ids_kmeans))
  cluster_memberships_kmeans_aligned[(cluster_memberships_kmeans==i)] <- max_cluster_overlap[i]

# Now calculate the overlaps
# First, the total overlap
total_observations_overlapping <- 100*sum(cluster_memberships_kmeans_aligned==cluster_memberships_hclust) / length(cluster_memberships_hclust)
# Then, per cluster
per_cluster_observations_overlapping <- sapply(1:length(cluster_ids_kmeans), function(i) 100*length(intersect(which(cluster_memberships_kmeans_aligned==i),which(cluster_memberships_hclust==i)))/sum(cluster_memberships_kmeans_aligned==i))
per_cluster_observations_overlapping <- matrix(per_cluster_observations_overlapping, nrow=1)
colnames(per_cluster_observations_overlapping) <- paste("Segment",1:length(per_cluster_observations_overlapping),sep=" ")
```

<br>
<br>

<div class="row">
<div class="col-md-3">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
print(xtable(per_cluster_observations_overlapping, caption = paste(paste("The percentage of observations belonging to the same segment is", total_observations_overlapping, sep=" "), "%."), digits=1), type="html", html.table.attributes = "class='table table-striped table-hover table-bordered'", caption.placement="top", comment = FALSE, include.rownames = FALSE)
```
</div>
</div>
<br>

How much overlap should we expect across clustering solutions?


--- 
## Key Technical Terms and Lessons

- Segmentation Variables
- Profiling Variables
- Distance Metrics
- Hierarchical Clustering
- Dendrogram
- K-means Clustering
- Robustness: Statistics, Interpretation, Decisions
- Actionability, Interpretability, Statistical Robustness

--- 
## Group Work

1. How many market segments (clusters) are there? Why?

2. How would you describe the segments?

3. How would the segments inform the strategy of CreeqBoat?


--- 

## Next Class: Discriminant Analysis, Purchase Drivers, and Linking Analytics to Business Decisions

