---
title: "Cluster Analysis and Segmentation: Interactive Tool"
author: "T. Evgeniou"
runtime: shiny
output: 
  html_document:
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r echo=FALSE, message=FALSE}
source("../../AnalyticsLibraries/library.R")

ggthemr('fresh')  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center")
options(knitr.kable.NA = '')

tags$style(type="text/css", "p { text-align:justify; }")
tags$style(type="text/css", "label { display: none; }")
tags$style(type="text/css", ".c3 svg, .dendroNetwork svg text { font-size:13px; font-family:\"Roboto\", \"Helvetica Neue\", Helvetica, Arial, sans-serif !important; }")
tags$style(type="text/css", ".formattable_widget { overflow:auto; max-height:425px; margin-bottom:23px; }")
tags$style(type="text/css", ".formattable_widget table { margin-bottom:0; }")
tags$style(type="text/css", ".formattable_widget td, .formattable_widget th { white-space: nowrap; }")

MAX_VARIABLES <- 20     # limits initial dataset filters
MAX_VARIABLES_PROFILE <- 100     # limits initial dataset filters
SAMPLE_ROWS <- 10       # limits sample dataset presentations
SAMPLE_XPOINTS <- 100   # limits sample dataset plots

normalize.abs <- function(x, min=0, max=1, na.rm=FALSE) normalize(abs(x), min, max, na.rm)
dformat <- function(df) {
  if (class(df) != "data.frame")
    df <- as.data.frame(df)
  x <- lapply(colnames(df), function(col) {
    if (is.numeric(df[, col]))
      color_bar(rgb(238, 238, 238, max=255), normalize.abs, min=0.1, na.rm=TRUE)
    else
      formatter("span")
  })
  names(x) <- colnames(df)
  formattable(df, x)
}
```


# Clustering and Segmentation in 9 steps

This tool follows the 9 steps for cluster analysis outlined in the [Cluster Analysis and Segmentation](http://inseaddataanalytics.github.io/INSEADAnalytics/Report_s45.html) reading of the course. 

First we load the data (`data/Mall_Visits.csv` by default):

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# Please ENTER the name of the file with the data used. The file should contain a matrix with one row per observation (e.g. person) and one column per attribute.

fileInput("dataFile", "",
          accept = c(
            "text/csv",
            "text/comma-separated-values,text/plain",
            ".csv"),
          width = "100%"
          )
ProjectDataX <- eventReactive(input$dataFile, ignoreNULL=FALSE, {
  datapath <- input$dataFile$datapath
  if (is.null(datapath))
    datapath <- "data/Mall_Visits.csv"
  df <- try(read.csv(datapath, stringsAsFactors=FALSE), silent=TRUE)
  df.valid <- class(df) == "data.frame" && any(sapply(df, is.numeric))
  if (!df.valid) {
    showModal(modalDialog(title="Error", HTML("The data file contains no valid numeric variables.<br><br>Please ensure it can be processed with `read.csv` in the command line. Common problems include a column separator other than the comma (`,`) and a decimal point other than the period (`.`).")))
    shiny::validate(need(df.valid, message=FALSE))
  }
  if (is.null(rownames(df)) || identical(rownames(df), as.character(1:nrow(df))))
    rownames(df) <- paste0("observation ", sprintf("%02i", 1:nrow(df)))
  df
})
```

We also need to select which variables to consider for the cluster analysis (`r MAX_VARIABLES` at most): 

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
selectizeInput("segmentationAttributes", "",
               choices = NULL,
               multiple = TRUE,
               options = list(
                 plugins = c("remove_button","drag_drop"),
                 maxItems = MAX_VARIABLES,
                 placeholder = "None"
               ),
               width="100%"
               )
observeEvent(ProjectDataX(), {
  attributes <- colnames(ProjectDataX())
  attributes <- attributes[sapply(ProjectDataX(), is.numeric)]
  updateSelectizeInput(session, "segmentationAttributes", choices=attributes, selected=head(attributes, MAX_VARIABLES))
})

ProjectData_segmentX <- eventReactive(input$segmentationAttributes, {
  ProjectDataX()[, input$segmentationAttributes, drop=F]
})
```

Here is how the first `r SAMPLE_ROWS` rows look like:

```{r}
renderFormattable(dformat(head(ProjectData_segmentX(), SAMPLE_ROWS)))
```


### Step 1: Confirm data is metric

The data we use here have the following descriptive statistics: 

```{r}
renderFormattable(dformat(round(my_summary(ProjectData_segmentX()), 2)))
```


## Step 2: Scale the  data

Select variables to standardize:

```{r}
# Please enter the variables you would like to standardize: 

selectizeInput("standardizedVariables", "",
               choices = NULL,
               multiple = TRUE,
               options = list(
                 plugins = c("remove_button","drag_drop"),
                 placeholder = "None"
               ),
               width="100%"
               )
observeEvent(ProjectData_segmentX(), {
  attributes <- colnames(ProjectData_segmentX())
  updateSelectizeInput(session, "standardizedVariables", choices=attributes, selected=attributes)
})

ProjectData_scaledX <- eventReactive(input$standardizedVariables, ignoreNULL=FALSE, {
  df <- ProjectData_segmentX()
  columns <- input$standardizedVariables
  if (length(columns) == 0)
    return(df)
  df[, columns] <- apply(df[, columns, drop=F], 2, function(r) {
    if (sd(r) != 0) 
      (r - mean(r)) / sd(r)
    else
      0*r
  })
  df
})

```

These are the summary statistics of the scaled dataset:

```{r}
renderFormattable(dformat(round(my_summary(ProjectData_scaledX()), 2)))
```


## Step 3: Select Segmentation Variables

We did this at the begining. 


## Step 4: Define similarity measure

We now select the distance metric eventually used for the clustering in case of hierarchical clustering (see `help(dist)`):

```{r}
# Please ENTER the distance metric eventually used for the clustering in case of hierarchical clustering (e.g. "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski" - see help(dist)). DEFAULT is "euclidean"
selectizeInput("distance_used", "",
               choices = c("euclidean", "maximum", "manhattan", "canberra", "binary", "minkowski"))
```


## Step 5: Visualize Pair-wise Distances 

This is the histogram of all pairwise distances for the `r renderText(input$distance_used)` distance:

```{r}
tags$div(style='height:480px', renderC3({
  Pairwise_Distances <- dist(ProjectData_scaledX(), method=input$distance_used)
  Pairwise_Histogram <- hist(Pairwise_Distances, plot=FALSE)
  c3(data.frame(x=head(Pairwise_Histogram$breaks, -1), Frequency=Pairwise_Histogram$counts), x="x", y="Frequency") %>% c3_bar(bar_width=0.90)
}))
```


## Step 6: Method and Number of Segments

Select the hierarchical clustering method to use (see `help(hclust)`)

```{r}
# Please ENTER the hierarchical clustering method to use (options are:
# "ward.D", "ward.D2", "single", "complete", "average", "mcquitty", "median" or "centroid")
# DEFAULT is "ward"
selectizeInput("hclust_method", "",
               choices = c("ward.D", "ward.D2", "single", "complete",
                           "average", "mcquitty", "median", "centroid"))

Hierarchical_ClusterX <- eventReactive({
  input$hclust_method
  input$distance_used
  ProjectData_scaledX()
},{
  Hierarchical_Cluster_distances <- dist(ProjectData_scaledX(), method=input$distance_used)
  Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=input$hclust_method)
  Hierarchical_Cluster
})
```

and the number of clusters to eventually use for analysis:

```{r}
# Please ENTER the number of clusters to eventually use for this report
sliderInput("numb_clusters_used", "", min=1, max=10, value=3)
```

This is the hierarchical clustering dendrogram:

```{r}
dendroNetworkOutput("cluster", height="480px")
output$cluster <- renderDendroNetwork({
  margins <- list(top=10, right=0, bottom=120, left=0)
  cluster <- Hierarchical_ClusterX()
  if (length(cluster$labels) > 40) cluster$labels <- NULL
  dendroNetwork(cluster, fontSize=14, treeOrientation="vertical", margins=margins, textRotate=90)
})
```

This is a plot of the "distances" traveled before we need to merge any of the lower and smaller in size clusters into larger ones - the heights of the tree branches that link the clusters as we traverse the tree from its leaves to its root. If we have `n` observations, this plot has `n-1` numbers:

```{r}
renderFormattable(dformat({
  h <- Hierarchical_ClusterX()$height
  df <- cbind(h[length(h):1], 1:length(h))
  colnames(df) <- c("distances","index")
  head(round(df, 2), SAMPLE_XPOINTS)
}))
```

We can also see the segment each observation (respondent in this case) belongs to for the first `r SAMPLE_ROWS` people:

```{r}
renderFormattable(dformat({
  cluster_memberships_hclust <- as.vector(cutree(Hierarchical_ClusterX(), k=input$numb_clusters_used)) # cut tree into 3 clusters
  cluster_ids_hclust=unique(cluster_memberships_hclust)

  ProjectData_with_hclust_membership <- cbind(1:length(cluster_memberships_hclust),cluster_memberships_hclust)
  colnames(ProjectData_with_hclust_membership)<-c("Observation Number","Cluster_Membership")
  round(head(ProjectData_with_hclust_membership, SAMPLE_ROWS), 2)
}))
```


### Using Kmean Clustering

Select the kmeans clustering method to use (see `help(kmeans)`):

```{r}
# Please ENTER the kmeans clustering method to use (options are:
# "Hartigan-Wong", "Lloyd", "Forgy", "MacQueen"
# DEFAULT is "Lloyd"
selectizeInput("kmeans_method", "",
               choices = c("Hartigan-Wong", "Lloyd", "Forgy", "MacQueen"),
               selected = "Lloyd")
```

Here are the clusters our observations belong to when we select `r renderText(input$numb_clusters_used)` clusters and the `r renderText(input$kmeans_method)` kmeans method, for the first `r SAMPLE_ROWS` data (note that the cluster IDs may differ from those from hierarchical clustering):

```{r}
renderFormattable(dformat({
  kmeans_clusters <- kmeans(ProjectData_scaledX(), centers=input$numb_clusters_used, iter.max=2000, algorithm=input$kmeans_method)

  ProjectData_with_kmeans_membership <- cbind(1:length(kmeans_clusters$cluster),kmeans_clusters$cluster)
  colnames(ProjectData_with_kmeans_membership)<-c("Observation Number","Cluster_Membership")
  rownames(ProjectData_with_kmeans_membership)<-NULL
  round(head(ProjectData_with_kmeans_membership, SAMPLE_ROWS), 2)
}))
```


## Step 7: Profile and interpret the segments 

We now select the original raw attributes to use for the profiling of the segments, the "profiling attributes" (`r MAX_VARIABLES_PROFILE` at most)

```{r}
# Please ENTER then original raw attributes to use for the profiling of the segments (the "profiling attributes")
selectizeInput("profile_attributes_used", "",
               choices = NULL,
               multiple = TRUE,
               options = list(
                 plugins = c("remove_button","drag_drop"),
                 maxItems = MAX_VARIABLES_PROFILE,
                 placeholder = "None"
               ),
               width="100%"
               )
observeEvent(ProjectDataX(), {
  attributes <- colnames(ProjectDataX())
  attributes <- attributes[sapply(ProjectDataX(), is.numeric)]
  updateSelectizeInput(session, "profile_attributes_used", choices=attributes, selected=head(attributes, MAX_VARIABLES))
})

ProjectData_profileX <- eventReactive(input$profile_attributes_used, {
  ProjectDataX()[, input$profile_attributes_used, drop=F]
})
```

and the method to use for profiling:

```{r}
# Please enter the method to use for the profiling (e.g. "hclust" or "kmeans"):
selectizeInput("profile_with", "",
               choices = c("hclust", "kmeans"))
```

```{r}
## TODO: THIS BECOMES INTERACTIVE

# Please enter the minimum number below which you would like not to print - this makes the readability of the tables easier. Default values are either 10e6 (to print everything) or 0.5. Try both to see the difference.
MIN_VALUE=0.5

```

The average values of our data for the total population as well as within each customer segment are:

```{r}
cluster.profileX <- eventReactive({
  input$profile_with
  ProjectData_profileX()
  input$kmeans_method
  input$numb_clusters_used
  input$hclust_method
  input$distance_used
},{
  cluster_memberships <- if (input$profile_with == "hclust"){
    Hierarchical_Cluster_distances <- dist(ProjectData_profileX(), method=input$distance_used)
    Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=input$hclust_method)
    as.vector(cutree(Hierarchical_Cluster, k=input$numb_clusters_used))

  } else { # kmeans
    kmeans_clusters <- kmeans(ProjectData_profileX(), centers=input$numb_clusters_used, iter.max=2000, algorithm=input$kmeans_method)
    kmeans_clusters$cluster
  }
  cluster_ids <- unique(cluster_memberships)

  # TODO: SAVE THE DATA in the cluster file
  # NewData = matrix(cluster_memberships,ncol=1)
  # write.csv(NewData,file=cluster_file)

  population_average <- matrix(apply(ProjectData_profileX(), 2, mean), ncol=1)
  colnames(population_average) <- "Population"
  Cluster_Profile_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_profileX()[(cluster_memberships==i), ], 2, mean))
  if (ncol(ProjectData_profileX()) <2)
    Cluster_Profile_mean=t(Cluster_Profile_mean)
  colnames(Cluster_Profile_mean) <- paste("Segment", 1:length(cluster_ids), sep=" ")
  cluster.profile <- cbind (population_average,Cluster_Profile_mean)
  cluster.profile
})

renderFormattable(dformat(round(cluster.profileX(),2)))
```

We can also "visualize" the segments using **snake plots** for each cluster. For example, we can plot the means of the profiling variables for each of our clusters to better visualize differences between segments. For better visualization we plot the standardized profiling variables.

```{r}
ProjectData_scaled_profileX <- eventReactive(ProjectData_profileX(), {
  df <- ProjectData_profileX()
  columns <- colnames(df)
  df[, columns] <- apply(df[, columns, drop=F], 2, function(r) {
    if (sd(r) != 0) 
      (r - mean(r)) / sd(r)
    else
      0*r
  })
  df
})

Cluster_Profile_standar_meanX <- eventReactive({
  input$profile_with
  ProjectData_scaled_profileX()
  input$kmeans_method
  input$numb_clusters_used
  input$hclust_method
  input$distance_used
},{
  cluster_memberships <- if (input$profile_with == "hclust"){
    Hierarchical_Cluster_distances <- dist(ProjectData_scaled_profileX(), method=input$distance_used)
    Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=input$hclust_method)
    as.vector(cutree(Hierarchical_Cluster, k=input$numb_clusters_used))

  } else { # kmeans
    kmeans_clusters <- kmeans(ProjectData_scaled_profileX(), centers=input$numb_clusters_used, iter.max=2000, algorithm=input$kmeans_method)
    kmeans_clusters$cluster
  }
  cluster_ids <- unique(cluster_memberships)

  Cluster_Profile_standar_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_scaled_profileX()[(cluster_memberships==i), ,drop = F], 2, mean))
  if (ncol(ProjectData_scaled_profileX()) < 2)
  Cluster_Profile_standar_mean = t(Cluster_Profile_standar_mean)
  colnames(Cluster_Profile_standar_mean) <- paste("Segment", 1:length(cluster_ids), sep=" ")
  Cluster_Profile_standar_mean
})

tags$div(style='height:480px', renderC3({
  c3(melt(as.data.frame(cbind(n=as.numeric(1:nrow(Cluster_Profile_standar_meanX())), Cluster_Profile_standar_meanX())), id="n"), x="n", y="value", group="variable") %>% c3_line('spline') %>%  xAxis(label="Profiling variables (standardized)") %>% yAxis(label="Mean of cluster")
}))
```

We can also compare the averages of the profiling variables of each segment relative to the average of the variables across the whole population. This can also help us better understand whether  there are indeed clusters in our data (e.g. if all segments are much like the overall population, there may be no segments). For example, we can measure the ratios of the average for each cluster to the average of the population minus 1 (e.g. `avg(cluster)` `/` `avg(population)` `-1`) and explore a matrix as the following one:

```{r}
cluster_profile_ratiosX <- eventReactive({
  input$profile_with
  ProjectData_profileX()
  input$kmeans_method
  input$numb_clusters_used
  input$hclust_method
  input$distance_used
},{
  cluster_memberships <- if (input$profile_with == "hclust"){
    Hierarchical_Cluster_distances <- dist(ProjectData_profileX(), method=input$distance_used)
    Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=input$hclust_method)
    as.vector(cutree(Hierarchical_Cluster, k=input$numb_clusters_used))

  } else { # kmeans
    kmeans_clusters <- kmeans(ProjectData_profileX(), centers=input$numb_clusters_used, iter.max=2000, algorithm=input$kmeans_method)
    kmeans_clusters$cluster
  }
  cluster_ids <- unique(cluster_memberships)

  population_average <- matrix(apply(ProjectData_profileX(), 2, mean), ncol=1)
  colnames(population_average) <- "Population"
  Cluster_Profile_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_profileX()[(cluster_memberships==i), ], 2, mean))
  if (ncol(ProjectData_profileX()) <2)
    Cluster_Profile_mean=t(Cluster_Profile_mean)
  colnames(Cluster_Profile_mean) <- paste("Segment", 1:length(cluster_ids), sep=" ")
  
  population_average_matrix <- population_average[,"Population",drop=F] %*% matrix(rep(1,ncol(Cluster_Profile_mean)),nrow=1)
  cluster_profile_ratios <- (ifelse(population_average_matrix==0, 0,Cluster_Profile_mean/population_average_matrix-1))
  colnames(cluster_profile_ratios) <- paste("Segment", 1:ncol(cluster_profile_ratios), sep=" ")
  rownames(cluster_profile_ratios) <- colnames(ProjectData_profileX())
  cluster_profile_ratios
})

renderFormattable(dformat(round(cluster_profile_ratiosX(), 2)))
```


## Step 8: Robustness Analysis

For the two `r renderText(input$numb_clusters_used)`-segments solutions found above (one using Kmeans and the other using Hierarchical Clustering), these percentages are as follows:

```{r}
per_cluster_observations_overlappingX <- eventReactive({
  ProjectData_profileX()
  input$kmeans_method
  input$numb_clusters_used
  input$hclust_method
  input$distance_used
},{
  Hierarchical_Cluster_distances <- dist(ProjectData_profileX(), method=input$distance_used)
  Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=input$hclust_method)
  cluster_memberships_hclust <- as.vector(cutree(Hierarchical_Cluster, k=input$numb_clusters_used))
  cluster_ids_hclust <- unique(cluster_memberships_hclust)

  kmeans_clusters <- kmeans(ProjectData_profileX(), centers=input$numb_clusters_used, iter.max=2000, algorithm=input$kmeans_method)
  cluster_memberships_kmeans <- kmeans_clusters$cluster
  cluster_ids_kmeans <- unique(cluster_memberships_kmeans)

# First, make sure the segment ids are correctly aligned
cluster_overlaps <- Reduce(cbind,lapply(1:length(cluster_ids_kmeans), function(i) {
  overlaps <- sapply(1:length(cluster_ids_hclust), function(j) {
    length(intersect(which(cluster_memberships_kmeans==i),
                     which(cluster_memberships_hclust==j))) } );
  overlaps}))
max_cluster_overlap = rep(0,length(cluster_ids_kmeans))
for (i in 1:length(cluster_ids_kmeans)){
  highest_now = which.max(cluster_overlaps)
  hclust_id_now = highest_now %% length(cluster_ids_kmeans)
  hclust_id_now = ifelse(hclust_id_now == 0, 3, hclust_id_now)
  kmeans_id_now = ceiling(highest_now/length(cluster_ids_kmeans))
  max_cluster_overlap[kmeans_id_now] <- hclust_id_now
  cluster_overlaps[hclust_id_now,] <- 0
  cluster_overlaps[,kmeans_id_now] <- 0
}
cluster_memberships_kmeans_aligned <- rep(0,length(cluster_memberships_kmeans))
for (i in 1:length(cluster_ids_kmeans))
  cluster_memberships_kmeans_aligned[(cluster_memberships_kmeans==i)] <- max_cluster_overlap[i]

# Now calculate the overlaps
# First, the total overlap
total_observations_overlapping <- 100*sum(cluster_memberships_kmeans_aligned==cluster_memberships_hclust) / length(cluster_memberships_hclust)
# Then, per cluster
per_cluster_observations_overlapping <- sapply(1:length(cluster_ids_kmeans), function(i) 100*length(intersect(which(cluster_memberships_kmeans_aligned==i),which(cluster_memberships_hclust==i)))/sum(cluster_memberships_kmeans_aligned==i))
per_cluster_observations_overlapping <- matrix(per_cluster_observations_overlapping, nrow=1)
colnames(per_cluster_observations_overlapping) <- paste("Segment",1:length(per_cluster_observations_overlapping),sep=" ")
per_cluster_observations_overlapping
})

renderFormattable(dformat(per_cluster_observations_overlappingX()))
```

<div style="height: 450px;" class="tocify-extend-page" data-unique="tocify-extend-page">
</div>
